### 2013-03-28 inst/examples/salmonport.Rmd
Substantial additions to salmonport.Rmd

% Portfolio prioritization of salmon conservation
% Sean Anderson

# Abstract #
First part is copied from what I wrote before (introduction and questions). Second part outlines the simulation. I haven't yet worked very much with the portfolios themselves, but I've set up the structure of the simulation. I outline how the simulation is currently working and show examples of some of the plotting functions I have to look at simulation and portfolio optimization.

# Introduction #
Managing risk is fundamental to the conservation of an endangered species. When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level. Typically we treat sources of risk at the metapopulation level as external and uncontrollable and so we manage risk by altering fishing or hunting on a population level as well as improving the connectivity of populations.

The management of financial portfolios provides another way of considering risk. Economists consider the risk and performance of a financial portfolio as a function of the weighting of individual assets that make up the portfolio. Modern Portfolio Theory (MPT) proposes that there is a set of portfolios that maximizes expected return for a level of expected risk or minimizes expected risk for a level of expected return [@Markowitz1952; @Markowitz1959]. This optimal set contains portfolios that range along a continuum of risk-tolerance; economists refer to this set as the efficient frontier.

Similarly, expected growth rate and variance at a metapopulation level is a function of the variance, covariance, and size of the individual populations. A portfolio approach to managing risk for a metapopulation therefore considers how conservation might affect the "weight" of each population in a metapopulation "portfolio".

Bring in salmon: we can manage risk at the population level or at the ESU level. ESU-level conservation is mandated but we generally lack tools to quantitatively manage risk at this level.

\begin{table}[h!]
\centering
\small
\caption{Components of salmon metapopulation portfolios}
\begin{tabular}{p{3.6cm}p{7.5cm}}
\toprule
Component & Definition for the salmon portfolio\\
\midrule
Assets & Viable salmonid populations? Stream-level populations? Depends on the 
system\\
Portfolio & The salmon metapopulation; possibly an ESU\\
Portfolio managers & Salmon managers\\
Investors & Salmon managers, conservation agency, or salmon fishers (each with 
their own goals and mandates)\\
Asset weights & Carrying capacity, or catches, or salmon spawner (or return) 
abundance\\
Asset returns & Annual rate of change of salmon returns or spawners; possibly 
annual rate of change of fisheries catches\\
Asset risk & Could be variance of asset returns; could be a coherent asymmetric 
risk metric such as CVaR\\
\bottomrule
\end{tabular}
\label{tab:port-components}
\end{table}

## Portfolio prioritization of salmon conservation ##
The basic goal is to conduct a post-hoc assessment of the set of optimal salmon portfolios given various return-risk definitions and objectives. We ask: What does a portfolio approach to management tell us about salmonid conservation priorities? Can we produce any rules of thumb about what optimal portfolio management of salmonid metapopulations would entail?

## Questions ##
1. What are the attributes of the populations that form the optimal set? For example, does the optimal set tend to focus on populations with as diverse responses to the environment as possible?

2. How much lower is stream-level and ESU-level extinction risk for the optimal portfolios than the less optimal portfolios? Extinction probability could be estimated through long-term simulation or through an asymmetric risk metric like conditional value at risk (CVaR). Given I am including straying, this will most likely be a quasi-extinction-risk metric.

3. How do our answers to other questions change depending on what we measure on the risk-return axes? For example, we could trade off the mean rate of change of salmon returns vs. variance of the rate of change. This would be most similar to a financial portfolio optimization. We could also trade off the salmon returns (or catches) themselves. Further, we could work with variance or a downside risk metric like CVaR.

4. How do constraints (e.g. viable salmonid population (VSP)-like constraints) affect the selected portfolios in risk-return space and in terms of common metrics such as population-level and ESU-level extinction risk?

5. How do the properties of the optimal portfolio set change based on different kinds of environmental driver patterns? For example, magnitude of driver variance, cyclical dynamics like the PDO, black swans or unexpected surprises, increases in temperature and variability through time (climate change).

# Simulation #
See Figure \ref{fig:sim-diagram} for an illustration of the simulation 
structure.

\begin{figure}[htbp]
  \centering
    \includegraphics[height=6.5in]{simulation_diagram.pdf}
 \caption{Flow chart of the salmon-metapopulation simulation and portfolio optimization. There are $n$ salmon populations and $t$ generations. Blue text indicates values that are generated before the simulation progresses through time. Orange text indicates steps in which calculations are performed. Black text indicates values that are calculated. Grey text indicates parameters that can be set. Green text indicates the looping structure of the simulation.}
  \label{fig:sim-diagram}
\end{figure}

## Spawner-return relationship ##
I am using a Ricker curve,

\begin{equation}
R_{ti} = S_{ti}e^{a_{ti}(1-S_{ti}/b_i) + w_{ti}},
\end{equation}

\noindent
where $t$ represents a generation time, $i$ represents a population, $R$ is the number of returns, $S$ is the number of spawners, $a$ is the productivity parameter (which can vary with the environmental signal), and $b$ is the density-dependent term (which is used as the asset weights in the portfolios). The term $w_{ti}$ represents first-order autocorrelated error (AR1). Formally, $w_{ti} = w_{ti-1} \rho_w + v_{ti}$, where $v_{ti}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_v$. The parameter $\rho_w$ represents the correlation between residuals from subsequent generations.

## Escapement targets ##
I am calculating spawning abundance MSY $S_{MSY}$ each generation using the equation $S_{MSY} = b(0.5-0.07a)$ from @Hilborn1992 p272, Table 7.2. I am using the precisely known $a$ and $b$ values each time and assume perfect implementation of the escapement goal (no outcome uncertainty) and no upper limit on how much the fishery can harvest. We may want to add a bit of noise here, say by assuming a constant value for $a$ or adding some implementation uncertainty. We need some level of harvesting to keep the spawning population away from carrying capacity. At carrying capacity, the productivity parameter $a$ has minimal effect on population dynamics, making it a poor way to integrate response diversity.

## Environmental signal ##
I have programmed five different kinds of environmental signals. I think it's easiest to think of these as temperature, but they could represent more general environmental anomalies, such as the PDO. In Table \ref{tab:env-types}, I describe the five kinds of environmental signals I have included. Examples of these time series are illustrated in Figure \ref{fig:env-ts}. Later, I may include black-swan-like events, although the regime shifts could be seen this way if the changes are dramatic enough. So far, all simulation results in the rest of this document use a sine wave. The problem with using a randomly generated ARMA time series is that you get much greater differences in the results from 
run to run. This is fine in the long run, but hard to detect patterns initially.

\begin{table}[h!]
\centering
\small
\caption{Environmental signals, parameters, and descriptions.}
\begin{tabular}{p{3.2cm}p{3.25cm}p{6.8cm}}
\toprule
Signal type & Parameters & Description \\
\midrule
Sine wave & $\overline{T} + A \cdot \sin(\omega t + \phi)$& $\overline{T}$ is 
the mean environmental value, $A$ is the amplitude, $\omega$ is the frequency in 
radians, and $\phi$ is the phase\\
ARMA & AR1, MA1, $\sigma_T$ & First-order autoregressive (AR1) and moving 
average (MA1) parameters and standard deviation; mean is assumed to be 0\\
Regime shifts & $T_{b0},T_{b1}, \ldots, T_{br-1}$; $T_{v0},T_{v1}, \ldots, T_{vr}$ 
& Breakpoints ($T_{b}$) and values ($T_v$) for regimes $1$ through $r$\\
Linear change & $T_{\mathrm{min}}$, $T_{\mathrm{max}}$ & Minimum and maximum 
values\\
Constant & $T_{\mathrm{constant}}$& Constant value\\
\bottomrule
\end{tabular}
\label{tab:env-types}
\end{table}

```{r env-ts, fig.height=4, fig.width=4, cache = FALSE, fig.cap="Example temperature time series\\label{fig:env-ts}"}
#library(metafolio)
#or for rapid development:
metafolio_files <-  list.files(path = "../../R/")
for(i in 1:length(metafolio_files)){
  source(paste("../..//R/", metafolio_files[i], sep = ""));rm(i)}

source("plot_env_ts.R")
col_pal <- rev(gg_color_hue(10))
```

## Straying ##
[Insert a bit here about my research on straying and it's influence on metapopulation dynamics for salmon]. Essentially, straying should be enough to prevent local extinctions and ensure all possible habitat is occupied, but probably not strong enough to noticeably affect population dynamics or synchrony [@Schtickzelle2007].

I've implemented straying as in @Cooper1999. I generate a matrix that represents the fraction of straying between any two populations. I have set up the metapopulation (thus far) in a very simple scenario: the populations are arranged in a line and those that are nearer to each other are more likely to stray between each other [insert caveats]. Two parameters control the straying: the fraction of fish $f_{stray}$ that stray from their natal stream in any given generation and the rate at which this straying between streams decays with distance $m$. I calculated the number of salmon straying from stream $j$ to stream $i$ as,

\begin{equation}
  \mathrm{strays}_{tij} = f_{stray} R_{tj} \frac{e^{-m \lvert i-j \rvert }}{\displaystyle\sum\limits_{\substack{k = 1 \\
  k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
\end{equation}

\noindent
where $R_{tj}$ is the number of returning salmon at generation $t$ whose natal stream was stream $j$. The subscript $k$ represents a stream ID and $n$ the number of populations. The denominator is a normalizing constant to ensure the desired fraction of fish stray. See Figure \ref{fig:stray-matrix} for an example straying matrix.

```{r stray-matrix, fig.height=3.7, fig.width=3.9, fig.cap="Example straying matrix. Darker blue colours indicate a greater rate of straying.\\label{fig:stray-matrix}"}
source("plot_straying_matrix.R")
```

## Diversity of response to the environmental conditions ##
[insert research I've done here on what can generate diversity of response to the environment across populations and how it is typically incorporated] This is a critical part to the paper.

I am incorporating diversity of response to environmental conditions, say temperature, via changes in the Ricker $a$ parameter. I am using thermal tolerance curves, based loosely on @Eliason2011. They identify unique thermal tolerance curves in terms of cardiorespiratory performance. I am using approximately the same temperature ranges and variability of those curves, but applying them to productivity. I show the curves I'm currently using in Figure \ref{fig:thermal-curves}. We could vary the distribution of the optimum temperatures, the maximum $a$ values, and the width of the curves if we wanted, although I think the current curves are a simple and straightforward base scenario and we may not need to do anything more complicated.

I calculate $a_{ti}$, the realized Ricker productivity parameters after accounting for thermal tolerance, as
\begin{equation}
  a_{ti} = a_i^* + \gamma_i (T_t - T_i^*)^2,
\end{equation}
\noindent
where $a_i^*$ is the maximum possible productivity parameter for population $i$, $T_t$ is the temperature at time step $t$, $T_i^*$ is the optimal temperature for population $i$, and $\gamma_i$ controls the width of the thermal tolerance curve for each population. See Figure \ref{fig:thermal-curves} for the default thermal tolerance curves for ten populations.

Another possibility would be to draw the $a^*$, $\gamma$, and $T^*$ randomly. The advantage to setting it up as I have (with pre-specified thermal tolerance curves) is that it will hopefully make the post-hoc analysis of what affects the efficient frontier weights more straightforward.

```{r thermal-curves, fig.height=3.7, cache = FALSE, fig.width=5.0, fig.cap="Thermal tolerance curves for ten populations. Here, and throughout this document, I have used the same rainbow of colours to identify the salmon populations. The populations are arranged from blue (more productive with cooler temperatures) to red (more productive with warmer temperatures). The vertical lines indicate mean, minimum, and maximum values for one scenario --- you can ignore them for now.\\label{fig:thermal-curves}"}
source("plot_thermal_curves.R")
```

# Summary of simulation parameters

\begin{table}[h!]
\centering
\small
\caption{Salmon metapopulation parameters with base case and alternate values.}
\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
\toprule
Description & Parameter & Base [lower, upper] \\
\midrule
Stock-recruit residual standard deviation (on log scale) & $\sigma_v$ & 0.30 [0.05, 0.50] \\
First order (AR1) serial correlation of stock-recruit residuals & $\rho_w$ & 0.40 [0, 0.80] \\
Fraction of fish that stray from natal streams & $f_{stray}$ & 0.02 [0, 0.10] \\
Exponential rate of decay of straying with distance & $m$ & 0.30 [0.05, 0.50] \\
Standard deviation of beta distribution for implementation error & $\sigma_{impl}$ & 0.05 [0, 0.20] \\
Frequency of assessment (years) & $f_{assess}$ & 20 [5, 50] \\
\bottomrule
\end{tabular}
\label{tab:salm-pars}
\end{table}

\clearpage

# Portfolio optimization #

Details to follow. Quick version: I Monte Carlo the $b$ parameters (with a specified minimum value) and measure the resulting mean and variance of the rate of growth of the metapopulation salmon returns. I then identify the set of $b$ parameters (i.e. the set of weights) that results in the efficient frontier. We could later choose to work with risk metrics like conditional value at risk (CVaR), which measure the probability of a ``bad'' event specifically, not the symmetrical variance.

To find the efficient portfolios I divide all portfolios into bins by mean rate of change and find the portfolios with the minimum variance within each bin. I discard all portfolios with a mean rate of change below the minimum variance portfolio. These are ``undesirable'' portfolios.

# Comparing portfolios #
This is a big area to think about. How do portfolios differ along the frontier? How do the frontier portfolios differ from non-frontier portfolios? What is the best way to visually compare them? What aspects to compare? Are there summary statistics we could use to compare them?

# Some sample simulations #

```{r run-base-example1, fig.height=6.5, fig.width=5, fig.cap="Example simulation with all parameters at default values. I'm only showing a sample of 40 generations (after the burn-in period). From top to bottom: salmon returns each generation (catch plus escapement); Ricker $a$ parameter; the environmental signal (could think of it as temperature or a composite environmental signal like the PDO); number of salmon straying by population; number of straying salmon being received by population; fisheries catch based on escapement MSY calculation; salmon escapement; residuals from the spawner-return curve; estimated Ricker $a$ parameter; estimated Ricker $b$ parameter."}
set.seed(2)
eg1 <- meta_sim(b = rep(1000, 10), n_pop = 10)
plot_sim_ts(eg1, years_to_show = 40)

#o <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.01, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
#o <- meta_sim(env_params = list(amplitude = 2.5, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.00001, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
#o <- meta_sim(env_params = list(amplitude = 2.2, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.1, n_pop = 10, stray_fraction = 0.01, use_cache = FALSE)
##o<- meta_sim(b = rep(1000, 10), n_pop = 10)
#plot_sim_ts(o, years_to_show = 40)
#image(o$Est_b)
```

```{r plot-base-corrs, fig.width=4, fig.height=4, fig.cap="A plot comparing the log(returns) between each population. The population colours and numbers match in all figures. Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics. Populations with thermal tolerance curves in the middle (e.g. population 6) are less correlated with other populations. Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity."}
plot_correlation_between_returns(eg1, pal = col_pal)
```

```{r plot-base-rickers, fig.width=6, fig.height=3.75, cache=FALSE, fig.cap="We can also look at samples of the Ricker curves by population (population numbers are in the upper left corners). Here, I've sampled 40 of the $a$ parameters through time for each population. The shaded rectangles to indicate the range of spawners observed in the simulation. The colour of the line indicates the temperature anomaly in that year with blue indicating cooler years and red indicating warmer years. In this example all the populations have the same $b$ parameters. Notice how the productivity is more variable for populations nearer the limits of their thermal tolerance (e.g. population 1 vs. population 5). Also notice how population 1 is more productive in cool years whereas population 10 is more productive in warm years."}
plot_rickers(eg1, pal = col_pal)
```

```{r plot-various-options-ts-3pops, fig.width=4, fig.height=6.3, fig.cap="Example simulations with different parameter values. I am showing three populations here to make the plots easier to interpret.\\label{fig:sim-param-ts-3}"}
source("example_returns_3.R")
```

\clearpage

# An example of portfolio optimization #

Now, let's run an example where we find the $b$ values that form the efficient frontier of metapopulation portfolios.

```{r set-a-seed, echo=FALSE}
set.seed(1)
```

```{r monte-carlo-portfolios}
weights_matrix <- create_asset_weights(n_pop = 10, n_sims = 100, 
  weight_lower_limit = 0.001)
mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
  n_sims = 100, mean_b = 1000)
```

```{r eff-frontier-eg, fig.width=6.5, fig.height=3.5, fig.cap="Efficient frontier of metapopulation portfolios (red dots). Each dot represents a different set of weights of the Ricker $b$ parameters. Metapopulation growth rate is the first difference of the log of total metapopulation abundance. The colours in the right panel correspond to the other figures."}
ef_dat <- plot_efficient_portfolios(port_vals = mc_ports$port_vals, pal = 
  col_pal, weights_matrix = weights_matrix)
```

```{r plot-eff-ports, fig.height=6.5, fig.width=5, fig.cap="Simulation output from the minimum-variance metapopulation."}
min_var_sim <- mc_ports$sims_out[[ef_dat$min_var_port_id]]
plot_sim_ts(min_var_sim, pal = col_pal)
```

```{r fig.width=6, fig.height=3.75, fig.cap="Example Ricker curves for the minimum-variance metapopulation."}
plot_rickers(min_var_sim, pal = col_pal)
```

```{r plot-portfolio-timeseries, fig.width=6, fig.height=3.5, fig.cap="Time series for a random sample of portfolios (grey lines) and the minimum-variance portfolio (white lines). The top panel shows rate of change or metapopulation abundance at each time step and the bottom panel shows the metapopulation abundance."}
source("plot_portfolio_ts.R")
```

```{r plot-weights-samples, fig.width=6, fig.height=6, fig.cap="A sample of 30 minimum-variance portfolios given the same set of starting parameters as for the example illustrated above. The dashed grey line indicates the weights if all populations had an equal carrying capacity. As it stands, the weights look fairly random to me."}
source("look_at_cluster_dat.R")
```

```{r plot-weights-median, fig.width=4.2, fig.height=3.2, cache = FALSE, fig.cap="Median (solid line) and 25th and 75th quantiles of the weights across 40 minimum-variance portfolios. Just another way of looking at it."}
source("min_var_ports_dist.R")
```

```{r plot-cons-plans-curves, cache=FALSE, fig.width=3.3, fig.height=5.0, fig.cap="Different ways of prioritizing conservation. The curves show the environmental-tolerance curves for the populations that are conserved. The colours are the same as earlier and identify the different populations. These are in the same order as they are referred to in the next two plots: full range of environmental response diversity, most stable only, one half, other half, most asynchronous.", cache=FALSE}
#Notes #
#-b is "unfished equilibrium stock size"
source("plot_thermal_curves_scenarios.R")
```

```{r plot-cons-plans-mv, fig.width=6.5, fig.height=3.5, fig.cap="Conservation plans in mean-variance space. The dots show portfolios from individual simulation runs. The dark and light shaded areas show 25% and 75% quartiles. The left panel shows a scenario with a stationary environmental signal and strong short-term environmental noise. The right panel shows a scenario with a linearly-increasing environmental signal and minimal short-term environmental noise. These two scenarios bound the two environmental extremes: strong short-term noise with minimal long-term signal vs. a strong long-term signal with minimal short-term signal. Notice what happens to the two *half* conservation strategies. A more realistic scenario combines the two. The efficient frontier is shown for the displayed portfolios. This isn't necessarily the overall efficient frontier.", cache=FALSE}
#Notes #
#-b is "unfished equilibrium stock size"
require(RColorBrewer)
cols <- brewer.pal(5, "Dark2")[c(1,2,4,5,3)] 
source("cons_plans.R")
```

```{r plot-cons-plans-ts, fig.width=6, fig.height=3, fig.cap="Conservation plans in terms of metapopulation abundance. One realization with a periodic (sine wave) environmental signal is shown as an example. The light-grey lines in the background show 100 randomly-sampled portfolios with the same stock-recruit residuals and the same environmental signal. Notice how much less variable the scenarios with balanced response diversity are. The two scenarios with unbalanced response diversity tend to seesaw with the environment."}
source("cons_plans_ts.R")
```

\clearpage

# Take home messages
1. If the dominant environmental signal is a long-term shift, and you keep around one half of the response diversity, then you get different portfolio means but the same variances.
2. If the dominant environmental signal is short-term, and you keep around one half of the response diversity, then you get different portfolio variances but the same means.
3. Combine those two (which best reflects reality) and you get both different means and variances.
4. Whether you invest in just stable, just asynchronous, or a mix of both, all that really matters is that you have as many populations as possible and as balanced as possible response diversity.
5. An even weighting across as many possible streams (assuming balanced response diversity) gets you pretty close to the minimum-variance portfolio. [not shown currently]
6. With systematic long-term environmental change that dominates: if you can forecast the environment and know the responses of salmon populations from individual streams then you can increase your expected rate of metapopulation growth with minimal effect on variance by prioritizing the populations that will do well. Although if you get it wrong you will do worse than if you had hedged your bets and conserved response diversity.
7. But, if there are also strong short-term environmental fluctuations (and there probably are), then there's always a benefit to portfolio variance by having some response diversity.
8. Variable stocks aren't inherently undesirable to conserve, as long as they are balanced or averaged out.
9. Take home message: "To keep every cog and wheel is the first precaution of intelligent tinkering." But, if you must prioritize, keep a selection of different cogs and
 wheels --- especially a spatially diverse set of cogs and wheels. That gives you the best chance of an efficient ecological portfolio.

# Open questions and next steps
1. How do smaller or larger values of the various parameters affect the results?
2. Are the results sensitive to the height and width of the thermal tolerance curves? (I think within reason, no.)
3. Are these results robust to, say, cyclical or regime shift style environmental noise?
4. Should this be a generic "environmental signal" or temperature?
5. Show what happens for a given scenario as you conserve more streams (it gets closer to the minimum-variance portfolio).
6. Show an example of where these conservation plans place you in mean-variance space across all possible weights. (Done but not shown.)
7. Bring these conservation plans back to established suggestions (e.g. VSP).
8. How can we simplify the main figures?
9. Is there a way to combine the main figures to make the result more understandable (e.g. little response-diversity curves as a legend)?
10. Redo the Monte Carlo portfolios to show an extreme example where you can see the weights shifting. For example, a linearly increasing environmental signal with some generation-to-generation noise.
11. Generations or years? Years seem simpler to talk about. Generations seem more widely applicable and accurate.

\clearpage

# References #