### 2013-04-09 inst/examples/salmportms.Rmd
copy sent to committee before 2013-04-08
updating whole project

% Portfolio prioritization of salmon conservation
% Sean C. Anderson and others
%

*Authors:*  
  Jonathan W. Moore  
  Trevor A. Branch  
  Andrew B. Cooper  
  Nicholas K. Dulvy  
  Maybe we should bring Michelle McClure back in

*Other title ideas:*  
- Portfolio optimization reveals...  
- something more general with animals instead of just salmon?  
- Portfolio theory reveals conservation strategies that...  
- Intelligent tinkering with ecological portfolios: patterns of environmental response diversity drive salmon conservation priorities  
- Keeping every cog and wheel...

  see Table A1 in "Viable Salmonid Populations and the Recovery of Evolutionarily Significant Units":
  straying rates from the literature for Pacific salmon. range from 0 to 90... usually around 4 or 5

# Introduction #

managing risk fundamental to conservation of metapopulations

portfolios provide another way of managing risk - explain mean-variance optimization and how this translates to ecology
[@Figge2004]
[@Hoekstra2012] 
[@Ando2011,@Ando2012] 
[@Markowitz1952,@Markowitz1959]

we use salmon as an example, describe the components of the example, provide references on metapopulation existence
[@Schindler2010]

research questions, and briefly our approach - Monte Carlo simulation modelling, as simple a model as possible while retaining important realistic aspects of salmon metapopulations

# Methods #

Data - if we're basing this at all on reality... probably not

## Simulation model components

general overview

## Salmon metapopulation dynamics ##

- spawner-recruit relationship paragraph
- straying paragraph

## Environmental dynamics ##

## Environmental response diversity ##
Scenarios? More variable are also more productive?...

## Portfolio optimization ##

## Conservation rules of thumb ##

# Results #

## Base case ##

## Effect of X on ... ##

## Effect of Y on ... ##

# Discussion #

# Acknowledgements #

# References #

# Tables #

1. Simulation parameters, symbols, base and alternate values

# Figures #

1. Flow chart of Monte Carlo procedure
2. Climate scenarios
3. Climate-productivity curves (response-diversity scenarios)
4. Time series components of an example simulation
5. Effect on returns of ramping up or down each main parameter
6. Efficient frontier given various climate scenarios
7. Rules of thumb (in the spirit of VSP guidelines) in risk-reward space
8. Potentially a plot that shows how some summary metric of the efficient frontier or rules of thumb for high and low scenarios of important simulation parameters

# Supplementary code #

The `metafolio` `R` package and documentation.

# Supplementary figures #

1. Straying matrix
2. Efficient frontiers and/or rules-of-thumb plots varying each major simulation variable up and down