### 2013-06-06 inst/ms/extra-param-table.tex
Move extra parameter table rows to temporary file

Target harvest rate                                              & 
$H_\mathrm{tar}$      & \\
Actual harvest rate                                              & $H_\mathrm{act}$     & \\
Salmon return abundance                & $R$ & \\
Salmon spawning abundance                & $S$ & \\
Ricker capacity parameter & $b$ & \\
Ricker productivity parameter & $a$ & \\
Environmental value at generation $t$ & $e_t$ & \\

Independent and normally distributed environmental deviations with mean 0 & $d$ & \\

Location and shape parameters for the implementation uncertainty beta 
distribution & $\alpha_h$, $\beta_h$ & \\
Mean of implementation uncertainty beta distribution & $\mu_h$ & \\

\midrule
Indices & & \\
\midrule
Stream population ID & $i$ and $j$ & \\
Generation & $t$ & \\
Number of populations & $n$ & \\