### 2013-06-11 inst/ms/ms.tex
Develop the main text

\emph{Other title ideas:}

\begin{itemize}
\item
  Prioritizing metapopulation conservation through the lens of portfolio
  theory
\item
  Portfolio optimization reveals\ldots{}
\item
  Portfolio theory reveals conservation strategies that\ldots{}
\item
  Intelligent tinkering with ecological portfolios: patterns of
  environmental response diversity drive salmon conservation priorities
\item
  Keeping every cog and wheel\ldots{}
\end{itemize}

\textbf{Ecology Letters}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  letters: 5000 words, 6 figures or tables
\item
  ideas and perspectives: 7500 words, 10 figures (needs 300 word
  proposal)
\end{itemize}

\begin{quote}
Ecology Letters is a forum for the very rapid publication of the most
novel research in ecology, research that is not yet in the public
domain. Manuscripts relating to the ecology of all taxa, in any biome
and geographic area will be considered, and priority will be given to
those papers exploring or testing clearly stated hypotheses. The journal
publishes concise papers that merit urgent publication by virtue of
their originality, general interest and their contribution to new
developments in ecology. We discourage purely descriptive papers and
those merely confirming or extending results of previous work.
\end{quote}

\textbf{Ecological Applications}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  60 pages including everything
\end{itemize}

\begin{quote}
Ecological Applications is concerned broadly with the applications of
ecological science to environmental problems. It publishes papers that
develop scientific principles to support environmental decision-making,
as well as papers that discuss the application of ecological concepts to
environmental issues, policy, and management. Papers may report on
experimental tests, actual applications, scientific decision support
techniques, economic analyses, social implications of environmental
issues, or other relevant topics. Statistical or experimental methods
papers that support research and applications are welcome. Papers
submitted to Ecological Applications should be accessible to both
scholars and practitioners.
\end{quote}

\section{Long and rambling abstract that's a bit old now}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Managing risk is fundamental to the conservation of an endangered
  species. When an endangered species exists as a metapopulation, we can
  manage risk at two levels: at the population level or at the
  metapopulation level. Whereas risk is typically managed at the
  population level, a portfolio approach to managing risk might consider
  how conservation affects the ``weight'' of each population in a
  metapopulation ``portfolio''.
\item
  Here, we ask how a portfolio approach to managing risk can inform the
  spatial conservation of metapopulations in a changing world. To answer
  this, we develop a salmon metapopulation simulation in which
  population-specific productivity is driven by spatially-distributed
  environmental tolerance and patterns of short- and long-term
  environmental change. We then implement different spatial conservation
  ``rules of thumb'' that control the population-specific carrying
  capacities and evaluate the salmon portfolios along risk and return
  axes, similarly to how financial portfolios are assessed.
\item
  Our results show, first, that maintaining populations with a variety
  of environmental tolerances gives the best chance at an efficient
  ecological portfolio --- minimizing metapopulation variance while
  maximizing metapopulation growth rate. This finding emphasizes the
  risk of allowing large spatial blocks of habitat destruction, say
  through the development of dams. Second, we show that focusing on
  well-performing stocks now at the detriment of others is at best
  equivalent to a risky but efficient portfolio and is more likely a
  risky and inefficient portfolio --- it neither minimizes
  metapopulation variance nor maximizes growth rate compared to other
  strategies. Third, we show that maintaining more populations reduces
  metapopulation risk for the same spatial conservation strategy. Given
  a lack of knowledge of how populations respond to the environment, the
  most risk-averse approach is to conserve as many populations as
  possible.
\item
  Our findings highlight three key points: (1) the conservation priority
  of maintaining biocomplexity and therefore environmental response
  diversity, (2) the research priority of identifying differences in
  environmental tolerance given predicted environmental changes, and (3)
  the utility of considering risk for groups of fish stocks ---
  especially given environmental, biological, and implementation
  uncertainty --- through the lens of portfolio theory.
\end{enumerate}

\section{Introduction}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Managing risk is fundamental to the conservation of an endangered
  species.

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    risk is usually assessed on a population-by-population basis
  \item
    for metapopulations, we can consider risk for components or risk for
    the whole
  \end{itemize}
\item
  The management of financial portfolios provides another way of
  considering risk for metapopulations.

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    performance of weighted assets
  \item
    efficient frontier concept
  \item
    the parallels with ecological metapopulations
  \end{itemize}
\item
  Pacific salmon are an ideal model to consider under the framework of
  portfolio prioritization because of their metapopulation structure,
  their societal importance and threatened status, and the need to
  prioritize conservation efforts.

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    portfolio-like structure

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      documented metapopulation structure
    \item
      predators and fisheries often integrate across stocks
    \end{itemize}
  \item
    Highly threatened; highly valued; face many threats

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      damns
    \item
      roads
    \item
      other habitat degradation
    \item
      logging
    \item
      fishing
    \item
      changing climate
    \end{itemize}
  \item
    There's a need for prioritization

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      because of the highly threatened nature
    \item
      conflicting goals for resource use
    \item
      limited resources
    \item
      Existing prioritization schemes
    \end{itemize}
  \end{itemize}
\item
  Portfolio stabilization through systematic asynchrony in salmon
  population dynamics can be driven by stocks responding uniquely to the
  same environment (`response diversity') or by physical features of the
  environment causing stocks to experience different environmental
  forces (`environmental filter').

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    different responses to the environment: response diversity /
    biocomplexity

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      genetic variation by stock
    \item
      can be driven through historical conditions
    \item
      we focus on this here
    \end{itemize}
  \item
    experiencing different environment: environmental filter concept
    (Moran like)

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      spatially spread out, geographical and hydrological features
      filter the environment to create a variety of forces
    \item
      Typically a combination of both
    \end{itemize}
  \item
    Conversely, forces acting against asynchrony

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      inverse: loss of population diversity
    \item
      inverse: moran
    \item
      but also: dispersal
    \item
      and: trophic interactions with synchronized species
    \end{itemize}
  \end{itemize}
\end{itemize}

Here, we ask how a portfolio approach to management can inform the
conservation of metapopulations in a changing world. We ask two primary
questions: (1) What does portfolio optimization tell us about different
spatial approaches to prioritizing metapopulation conservation assuming
that response diversity is spatially distributed? (2) If we don't know
how response diversity is distribute, what does portfolio optimization
tell us about how many populations we should conserve? To answer these
questions, we develop a salmon metapopulation simulation in which
population-specific productivity is driven by spatially-distributed
environmental tolerance and patterns of short- and long-term climatic
change. We then implement different conservation ``rules of thumb'' that
control the population-level carrying capacities and evaluate the salmon
portfolios along risk and return axes, as a portfolio manager would. Our
findings illustrate that: (1) conserving response diversity buffers
metapopulation risk given short-term climate forcing and rate of return
given long-term climate forcing, (2) conserving more subpopulations
buffers risk regardless of the climate trend, and (3) considering
metapopulations through the lens of portfolio theory provides a useful
additional dimension through which we can evaluate conservation
strategies.

\section{Methods}

We developed a salmon metapopulation simulation model that includes a
stock-recruit relationship, demographic stochasticity, straying between
populations, varying responses to the environment, escapement target
setting, and implementation uncertainty. Under two kinds of
environmental regimes we tested different conservation rules of thumb
and evaluated these plans in risk-return space similar to how financial
managers evaluate financial portfolios. See Figure \ref{f:sim-flow} for
an illustration of the overall simulation structure. See Table
\ref{t:pars} for a list of the input parameters to our simulation and
their default values. We provide a package \texttt{metafolio} for the
statistical software \texttt{R} \citep{r2013} as an appendix, to carry
out the simulations and analyses we describe in this paper.

\subsection{Defining the ecological portfolio}

In our ecological portfolios, we defined assets as stream-level
populations and the portfolios as salmon metapopulations (Table
\ref{t:port}). We use the terms \textit{stream} and \textit{populations}
interchangeably to represent the portfolio assets. We defined the
portfolio investors as the stakeholders in the fishery and
metapopulation performance. For example, we could consider fisheries
managers, conservation agencies, or First Nations groups as investors.
We defined asset value as the abundance of returning salmon in each
stream and value of the portfolio as the overall metapopulation
abundance. In this scenario, the equivalent to financial rate of return
is the generation-to-generation rate of change of metapopulation
abundance. We defined the financial asset investment weights as the
capacity of the stream populations --- specifically the unfinished
equilibrium stock size --- since maintaining or restoring habitat
requires money, time, and resources. Investment in a population
therefore represents investing in salmon habitat conservation or
reconstruction.

\subsection{Salmon metapopulation dynamics}

The salmon metapopulation dynamics in our simulation were governed by a
spawner-return relationship with demographic stochasticity and by
straying between populations.

\subsubsection{Spawner-return relationship}

We defined the spawner-return relationship with a Ricker model
\citep{ricker1954},

\begin{equation}
R_{ti} = S_{ti}e^{a_{ti}(1-S_{ti}/b_i) + w_{ti}}
\end{equation}

\noindent where $t$ represents a generation time, $i$ represents a
population, $R$ is the number of returns, $S$ is the number of spawners,
$a$ is the productivity parameter (which can vary with the environment),
and $b$ is the density-dependent term (which is used as the asset
weights in the portfolios). The term $w_{ti}$ represents first-order
autocorrelated error. Formally, $w_{ti} = w_{ti-1} \rho_w + r_{ti}$,
where $r_{ti}$ represents independent and normally distributed error
with mean 0 and standard deviation of $\sigma_r$. The parameter $\rho_w$
represents the correlation between residuals from subsequent
generations.

We manipulated the capacity and productivity parameters $b$ and $a$ as
part of the portfolio simulation. The capacity parameters $b_i$ were
controlled by the investment weights in the populations. For example, a
large investment in a stream was represented by a larger unfished
equilibrium stock size $b$ for stream $i$. The productivity parameters
$a_{ti}$ were controlled by the interaction between an environmental
signal and the stream-level population environmental-tolerance curves.

We generated the environmental-tolerance parabolas according to

\begin{equation}
  a_{ti} =
  \begin{cases}
    W_i (e_t - e_i^{\mathrm{opt}})^2 + a_i^{\mathrm{max}},
      & \text{if } a_{ti} > 0\\
      0, & \text{if } a_{ti} \leq 0
  \end{cases}
\end{equation}

\noindent where $W_i$ controls the width of the curve for population
$i$, $e_t$ represents the environmental value at generation $t$,
$e_i^{\mathrm{opt}}$ represents the optimal environmental value for
population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible
$a$ value for population $i$. We set the $W_i$ parameters (Table
\ref{t:pars}) and calculated the $a_i^{\mathrm{max}}$ parameters so that
the area under each curve $A_i$ was equal. See Figure \ref{f:curves}a
for example environmental-tolerance curves.

\subsubsection{Straying}

We implemented straying as in \citet{cooper1999}. We set up the
metapopulation in a simple scenario: we arranged the populations in a
line and those that were nearer to each other were more likely to stray
between each other. Two parameters controlled the straying: the fraction
of fish $f_{\mathrm{stray}}$ that stray from their natal stream in any
generation and the rate $m$ at which this straying between streams
decays with distance. We calculated the number of salmon straying from
stream $j$ to stream $i$ as

\begin{equation}
  \mathrm{strays}_{tij} = f_{\mathrm{stray}} R_{tj}
    \frac{e^{-m \lvert i-j \rvert }}
      {\displaystyle\sum\limits_{\substack{k = 1 \\
    k \neq j}}^{n} e^{-m \lvert k-j \rvert }}
  \label{eq:stray}
\end{equation}

\noindent where $R_{tj}$ is the number of returning salmon at generation
$t$ whose natal stream was stream $j$. The subscript $k$ represents a
stream ID and $n$ the number of populations. The denominator is a
normalizing constant to ensure the desired fraction of fish stray. See
Figure \ref{f:stray} for an example straying matrix.

\subsection{Fishing}

Our simulation used a simple set of rules to establish escapement
targets and harvest the fish. Every $f_\mathrm{assess}$ years (default
of five years) our simulation fitted a spawner-return function and
target harvest rate $H_{\mathrm{tar}}$ was set based on
\citet{hilborn1992} as

\begin{equation}
  H_{\mathrm{tar}} = \frac{A}{b (0.5 - 0.07a)}
  \label{eq:esc}
\end{equation}

\noindent where $A$ represents the return abundance and $a$ and $b$
represent the Ricker model parameters. We included implementation
uncertainty in the actual harvest rate $H_{\mathrm{act}}$ as

\begin{equation}
  H_{\mathrm{act}} = \mathrm{beta}(\alpha_h, \beta_h)
\end{equation}

\noindent where $\alpha_h$ and $\beta_h$ are the location and shape
parameters in a beta distribution. They can be calculated from the
desired mean $\mu_h$ and standard deviation $\sigma_h$ as
\citep[p.~97]{morgan1990}

\begin{align}
  \alpha_h &= \mu_h^2
                \left(
                \frac{1 - \mu_h}{\sigma_h^2} - \frac{1}{\mu_h}
                \right)\\
   \beta_h &= \alpha \left({\frac{1}{\mu_h} - 1}\right).
\end{align}

\noindent Further, to establish a range of spawner-return values and to
mimic the start of an open-access fishery, for the first 30 years we
drew the fraction of fish harvested randomly from a uniform distribution
between 0.1 and 0.9. We discarded these initial 30 years as a burn-in
period throughout our analyses.

\subsection{Environmental dynamics}

We evaluated portfolio performance under short- and long-term
environmental dynamics. We represented short-term dynamics as a
stationary first-order autoregressive process, AR(1), with correlation
$\rho_e$

\begin{equation}
  e_t = e_{t-1} \rho_e + d_t, d_t \sim \mathrm{N}(0, \sigma_d)
\end{equation}

\noindent where $e_t$ represents the environmental value in generation
$t$ and $d$ represents normally distributed deviations of mean 0 and
standard deviation $\sigma_d$. We represented long-term environmental
dynamics as a linear shift in the environmental value through time

\begin{equation}
  e_t = \beta_e t - \overline{\beta_e t}
\end{equation}

\noindent where $\beta_e$ represents the slope. To maintain a balanced
response, we centered the trend by subtracting the mean
$\overline{\beta_e t}$ so that midway through the simulation (after any
burn-in period) the environmental value was at the mean environmental
tolerance.

\subsection{Conservation rules of thumb}

We evaluated two sets of conservation rules of thumb: (1) spatial
response diversity conservation strategies in an idealistic scenario
where you can detect response diversity, and (2) a more realistic
scenario where we know little about response diversity and we're left
with a choice of how many populations to conserve.

We evaluated four spatial conservation rules of thumb (Figure
\ref{f:curves}b--e). In all spatial scenarios, we conserved four
populations and set the unfished equilibrium biomass of the remaining
populations to near elimination (five salmon). These reduced populations
could still receive straying salmon but were unlikely to rebuild on
their own to a substantial abundance. The four scenarios were:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Conserve an even sampling of response diversity.
\item
  Conserve the most stable populations only.
\item
  Conserve one half of the metapopulation.
\item
  Conserve the other half of the metapopulation.
\end{enumerate}

In reality we rarely know precise levels of response diversity. We
therefore additionally considered a case where the conservation was
randomly assigned with respect to response diversity but where different
numbers of streams could be conserved. We considered conserving from two
to 16 streams. Similarly to the spatial strategies, we reduced the
capacity of the remaining streams to the nominal level of five salmon.

\section{Results}

\subsection{Which populations to conserve?}

\subsubsection{Short-term environment}

Given strong short-term environmental fluctuations, conserving response
diversity buffers the risk properties of an ecological portfolio (Fig.
4a). In our simulation, the median variance of generation-to-generation
rate of change in abundance was X times lower given balanced response
diversity (full range of responses or most stable only vs.~conserving
one half or the other). In fact, even though by conserving the full
range of responses, the portfolio was comprised of warm and
cool-thriving populations that were more variable on their own, each was
balanced by an opposing population. The portfolio risk was therefore
comparable between the full range of responses and most stable only
portfolios.

We can see the mechanism behind these portfolio properties by inspecting
example population time series (Fig. 4c, d). If only the upper or lower
half of response diversity is conserved, the portfolio tends to do well
or poorly depending on the environmental conditions (Fig. 4d). This risk
is buffered with balanced response diversity (Fig. 4c).

\subsubsection{Long-term environment}

Given long-term environmental change, the choice of which populations to
conserve affects the return properties of an ecological portfolio (Fig.
4b). By conserving balance response diversity, an ecological manager is
hedging his or her bets on what will happen with the environment and how
the populations will respond. The typical return for a balanced response
diversity strategy was zero --- the metapopulation neither increased or
decreased in abundance in the long run. By conserving only the upper or
lower half of response diversity, a conservation manager is putting all
his or her eggs in one basket --- the metapopulation might do really
well through time or it might do really poorly. The example
metapopulation abundance time series (Fig. 4d, f) illustrate this
effect. By conserving response diversity, when one population is doing
poorly, another is doing well and the metapopulation abundance remains
stationary through time.

Notably, in theses simulations, if a managers invested in the
populations that were doing well at the beginning they would have had
the lowest rate-of-return portfolio in the end (purple portfolios in
Fig. 4b).

Spatial conservation strategies in the face of longterm environmental
change hinge on whether you ``get it right'' --- whether you choose just
the right populations to conserve.

\subsection{How many populations to conserve?}

Given a scenario where we don't know the distribution of
population-level response diversity, portfolio optimization informs us
about the risk buffering from maintaining multiple populations (Fig. 5).
In short, investing in more populations buffers portfolio risk.

\subsubsection{Short-term environment}

Given short-term environmental noise, conserving more populations
buffers portfolio risk while the random conservation of response
diversity creates a spread of metapopulation risk for the same number of
populations conserved (Fig. 5a). For example, a metapopulation with
eight conserved populations is X times less risky than a metapopulation
with only four. We can see this risk-buffering effect through example
metapopulations in Fig. 5c and 5d. We note that the risk-return axes of
portfolio optimization ignore the absolute-abundance dimension (Fig.
5d). As one would expect, conserving fewer populations also results in
lower-abundance metapopulations.

\subsubsection{Long-term environment}

Given long-term environmental noise, conserving more populations also
buffers portfolio risk. However, in contrast to the short-term
environmental noise scenario, the unknown response diversity creates a
spread of possible metapopulation return for the same number of
conserved populations (Fig. 5b). Here, the number of populations
conserved buffers non-systematic (i.e.~not-environmentally-driven)
stochasticity.

\section{Discussion}

\section{Acknowledgements}

\bibliographystyle{apalike}

\bibliography{jshort,ms}

\clearpage

\section{Tables}

\begin{table}[h!]
\centering
\small
\caption{Components of salmon metapopulation portfolios KEEP THIS?}
\begin{tabular}{p{3.6cm}p{7.5cm}}
\toprule
Component          & Definition for the salmon portfolio\\
\midrule
Assets             & Stream-level salmon populations; possibly a Viable Salmonid Population\\
Portfolio          & The salmon metapopulation; possibly an Evolutionarily Significant Unit\\
Portfolio managers & Salmon managers\\
Investors          & Salmon managers, conservation agency, or salmon fishers\\
Asset weights      & Carrying capacity (specifically the $b$ parameters in a Ricker model)\\
Asset returns      & Rate of change of generation-to-generation salmon metapopulation abundance\\
Asset risk         & Variance of generation-to-generation salmon metapopulation abundance\\
\bottomrule
\end{tabular}
\label{t:port}
\end{table}

\clearpage

\begin{table}[h!]
\centering
\footnotesize
\caption{Input parameters to the salmon metapopulation simulation with default values.}
\begin{tabular}{p{7.7cm}p{1.4cm}p{3.6cm}}
\toprule
Description                                                          & Symbol                & Value \\
\midrule

\textit{Population dynamics parameters}                              &                       & \\
Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.30  \\
AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.40  \\
Fraction of fish that stray from natal streams                       & $f_{\mathrm{stray}}$  & 0.02  \\
Exponential rate of decay of straying with distance                  & $m$                   & 0.3  \\

\noalign{\vskip 3mm}
\textit{Environmental parameters}                                    &                       & \\
Width of environmental-tolerance curves for populations $i$ 1 to $n$ & $W_i$                 & seq(0.05, 0.02, 0.02 0.05, length = n pop)\\
Optimum environmental value for populations $i$ 1 to $n$             & $e_i^{\mathrm{opt}}$  & seq(13, 19, length = n pop)\\
Area under each environmental-tolerance curve in environmental units & $A$                   & 30\\

Standard deviation of short-term environmental fluctuations          & $\sigma_d$            & 2 \\
AR(1) autocorrelation of short-term environmental fluctuations       & $\rho_e$              & 0.1 \\
Slope of long-term environmental signal                              & $\beta_e$             & 0.114 \\

\noalign{\vskip 3mm}
\textit{Fishery parameters}                                          &                       & \\
Standard deviation of beta distribution for implementation error     & $\sigma_{h}$          & 0.05  \\
Frequency of assessment (years)                                      & $f_{\mathrm{assess}}$ & 5  \\
\bottomrule
\end{tabular}
\label{t:pars}
\end{table}

\clearpage

\section{Figures}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[height=5.5in]{../examples/simulation-diagram2.pdf}
\caption{Flow chart of the salmon-metapopulation simulation. There are $n$ salmon populations and $t$ generations. Blue text indicates values that are generated before the simulation progresses through time. Red text indicates steps in which calculations are performed through time. Black text indicates values that are calculated. Grey text indicates parameters that can be set. Green text indicates the looping structure of the simulation.}
\label{f:sim-flow}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
\caption{Different ways of prioritizing response-diversity conservation. Panel a shows the thermal tolerance cures for ten possible populations and panels b--e show different ways of prioritizing four of those populations. The curves describe how productivity varies with the environment for a given population. Some populations thrive at low environmental values (cool colours) and some at high (warm colours) values. Some are tolerant to a wider range of environmental conditions (yellow-to-green colours) but with a lower maximum productivity. The total possible productivity (the area under the curves) is the same for each population.}
\label{f:curves}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.0in]{../examples/spatial-arma-sim.pdf}
\caption{The components of an example metapopulation simulation.  We show, from top to bottom, the environmental signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, stock-recruit residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve. The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) environmental values.}
\label{f:sp-eg}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=5.0in]{../examples/spatial-mv.pdf}
\caption{The importance of preserving environmental response diversity through spatial conservation strategies. The conservation strategies correspond to Fig.~\ref{f:curves} and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).  In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.~X). The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy. We also show example metapopulation abundance time series for the (c, e) short-term and (d, f) long-term  environmental-fluctuation scenarios.}
\label{f:sp-mv}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=5.0in]{../examples/cons-plans-n.pdf}
\caption{The importance of preserving as many subpopulations as possible when we don't know how response diversity is distributed. In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.~X). We show metapopulations in which 2 (red), 4 (orange), 8 (yellow), or 16 (green) populations of random response diversity are conserved. The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy. We also show example metapopulation (c) rate-of-change and (d) abundance time series for the short-term environmental-fluctuation scenario.}
\label{f:n-mv}
\end{figure}

\clearpage