### 2013-08-26 inst/ms/salmonfolios.tex

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[round]{natbib}
\bibpunct{(}{)}{;}{a}{}{;}
%\textheight 20.5cm
%\topmargin -1.43cm
%\oddsidemargin 1.2cm
%\evensidemargin 1.2cm
%\textwidth 14.5cm
\usepackage{lineno}
\usepackage{xcolor}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}

\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}

\usepackage{fontspec}
\defaultfontfeatures{Ligatures=TeX}
\setsansfont{Calibri}
\setmonofont{Inconsolata}
\setmainfont{Minion Pro}

% remove numbers in front of sections:
\makeatletter
\renewcommand\@seccntformat[1]{}
\makeatother

\hyphenation{meta-pop-ulation meta-pop-ulations sub-pop-ulations sub-pop-ulation e-con-o-mist en-vi-ron-men-tal pri-or-i-ti-za-tion}

\begin{document}

\linenumbers
\modulolinenumbers[2]
\begin{spacing}{1.4}
\setlength{\parskip}{12pt}
\setlength{\parindent}{0cm} 
\include{ms}

% redefine the command that creates the Table number.
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0} % reset counter

% redefine the command that creates the Figure number.
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\figurename}{Fig.}
\setcounter{figure}{0}  % reset counter

\include{som}

\end{spacing}
\end{document}