### 2013-09-04 inst/ms/som.tex
Make a variety of edits to ms; finish SOM figs

\section{Supplementary Information}

\subsection{Supplementary code}

The \texttt{metafolio} \texttt{R} package and documentation. Some
details on what you can do with the package.

\subsection{Supplementary tables}

\input{tables}

\subsection{Supplementary figures}

\begin{figure}[htbp]
\centering
\includegraphics[width=3.5in]{../examples/figure/stray-matrix.pdf}
\caption{An example straying matrix. The rows and columns represent different 
populations (indicated by population number). Dark blue indicates a high rate 
of straying and light blue indicates a low rate of straying.}
\label{f:stray}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.0in]{../examples/plot-various-options-ts-3pops.pdf}
\caption{The impact of increasing or decreasing various parameter values on metapopulation return abundance. The different coloured lines represent three example salmon populations. The base case represents the base-case values for the short-term environmental fluctuation scenario.}
\label{f:eg-sens}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/spatial-arma-sim-full.pdf}
\caption{Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations.}
\label{f:eg-sp-arma}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/spatial-arma-sim-onehalf.pdf}
\caption{Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations.}
\label{f:eg-sp-arma}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/spatial-linear-sim-full.pdf}
\caption{Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental fluctuations.}
\label{f:eg-sp-linear}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/spatial-linear-sim-onehalf.pdf}
\caption{Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental fluctuations.}
\label{f:eg-sp-linear}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/n-arma-sim-2.pdf}
\caption{\textbf{Two populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.}
\label{f:eg-n-arma}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/n-arma-sim-16.pdf}
\caption{\textbf{Sixteen populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.}
\label{f:eg-n-arma}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/n-linear-sim-2.pdf}
\caption{\textbf{Two populations} conserved with random response diversity and \textbf{long-term} environmental change.}
\label{f:eg-n-linear}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/n-linear-sim-16.pdf}
\caption{\textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term} environmental change.}
\label{f:eg-n-linear}
\end{figure}