### 2013-09-04 vignette/metafolio.Rnw
Add in-progress vignette

%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{metafolio}

% tw=68
\documentclass[12pt]{article}
\usepackage{geometry}
\usepackage{url}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}

\setlength{\parskip}{12pt}
\setlength{\parindent}{0cm}

\begin{document}

<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
# set global chunk options
opts_chunk$set(fig.path='figure/', fig.align='center', fig.show='hold')
options(replace.assign=TRUE, width=90)
opts_knit$set(out.format = "latex")
opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, tidy=FALSE, 
refresh=TRUE)
#knit_theme$set("solarized-light")
@

\title{A guide to using the \texttt{metafolio} \texttt{R} package}
\author{Sean C. Anderson, and co-authors}
\date{}
\maketitle

The \texttt{metafolio} \texttt{R} package is a tool to simulate salmon
metapopulations and apply financial portfolio optimization concepts
to those metapopulations. This vignette accompanies the paper
\textit{Portfolio conservation of metapopulations under climate
change}. In this vignette we will describe the main functions in the
\texttt{metafolio} package, show how to re-create the main analyses
in our paper, and demonstrate some of the included plotting
functions for exploring the output.

\section{Installing the package}

To install and load the package:

<<my-load-package, include=FALSE>>=
library(metafolio)
set.seed(1)
@

<<load-package, eval=FALSE>>=
install.packages("devtools") # if needed
devtools::install_github("metafolio", "seananderson")
library(metafolio)
@

You can get more detailed installation instructions, read the source
code, and report bugs on GitHub:
\url{https://github.com/seananderson/metafolio}

\section{An example simulation}

The main simulation function is \texttt{meta\_sim}. We'll start by
running a simulation using the base-case parameter values from the
paper. First, we'll set up a list object that contains the argument
values for the stationary environmental stochasticity scenario. The
arguments \texttt{mean\_value}, \texttt{ar}, and \texttt{ma} refer
to the mean, autoregressive (AR), and moving-average (MA) components
of an ARIMA model. See \texttt{?arima} for further explanation of
ARIMA models in R. The argument \texttt{sigma\_env} refers to the
standard deviation.

<<setup-arma>>=
arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, 
  ma = 0) 
@

Next we'll run the simulation for one iteration. We'll simulate ten
subpopulations and we'll re-assess the fishery every five years.
Re-assessing the fishery means that the simulation fits a Ricker
curve to the observed spawner-return data and updates the
harvest-level targets.

<<base-case-run, cache=TRUE>>=
base1 <- meta_sim(n_pop = 10, env_params = arma_env_params, 
  env_type = "arma", assess_freq = 5)
@

We can plot the output with the function \texttt{plot\_sim\_ts}. The
output is shown in Figure~\ref{fig:plot-base-case-ts}.

<<plot-base-case-ts, fig.width=5, fig.height=6.5, out.width="4in", fig.cap="An example simulation with stationary environmental stochasticity and the base-case parameter values.">>=
plot_sim_ts(base1, years_to_show = 70, burn = 30)
@

\clearpage

\section{Exploring prioritization strategies}

One of the key elements to the analysis in our paper is the choice
of which populations to prioritize for conservation. We can try
different prioritization strategies by manipulating the "investment
weights" in each stream of salmon. In the case of salmon, we
represent these with the Ricker $b_i$ parameters, which represent
the maximum population capacity of streams $i$ 1 through $n$.

In this example, we'll create a scenario in which we conserve a
variety of balanced response diversity and a scenario in which we
conserve one half of the response diversity. We've set up the
weights carefully so that each metapopulation contains the same
number of subpopulations (10) and the same total capacity. We set
$b$ equal to the nominal level of 5 for the streams that we choose
not to prioritize.

<<spatial-plans>>=
w_plans <- list()
w_plans[["balanced"]] <- matrix(c(5, 1000, 5, 1000, 5, 5, 1000, 5, 
    1000, 5), nrow = 1) 
w_plans[["one_half"]] <- c(rep(1000, 4), rep(5, 6))
w <- list()
for(i in 1:2) { # loop over plans
 w[[i]] <- list()
 for(j in 1:80) { # loop over iterations
   w[[i]][[j]] <- matrix(w_plans[[i]], nrow = 1)
 }
}
@

We've now created a nested list object of stream capacities
(\texttt{w}). The first level of the list contains the two different
plans. The second level of the list contains the $b$ values for each
iteration. Each iteration will have re-sampled process noise and
observation error when run with \texttt{meta\_sim}. Here, we're
keeping the $b$ values the same between iterations, but that might
not be the case if we wanted to stochastically simulate the $b$
values. We're only going to run 80 iterations to reduce the runtime
of the example, but in reality you would likely want to run many
more iterations.

We can now stochastically simulate with these strategies using the
function\\ \texttt{run\_cons\_plans} and plot the output with
\texttt{plot\_cons\_plans} (Figure~\ref{fig:spatial-runs}).

<<spatial-runs, fig.width=5, fig.height=5, out.width="4in", cache=TRUE, fig.cap="Two spatial conservation strategies shown in risk-return space with stationary environmental stochasticity. The dots show simulated metapopulations and the contours show 25\\% and 75\\% quantiles across 80 simulations per strategy. The grey line indicates the efficient frontier across all simulated metapopulations.">>=
set.seed(1)
arma_sp <- run_cons_plans(w, env_type = "arma", env_params =
  arma_env_params)
plot_cons_plans(arma_sp$plans_mv, 
  plans_name = c("Balanced", "One half"), 
  cols = c("#E41A1C", "#377EB8"), xlab = "Variance of growth rate", 
  ylab = "Mean growth rate")
@

\section{Generating alternative environmental time series}

We can use the function \texttt{generate\_env\_ts} to create
alternative environmental time series. The function can generate
five kinds of time series: sine waves, regime shifts, linear
changes, and constant values. Each type has its own set of parameter
arguments that are passed in a list format. See
\texttt{?generate\_env\_ts} for examples of all of these.

We can see demonstrations of all environmental time series types
with the default argument values with the following code
(Figure~\ref{fig:env-eg}).

<<env-eg, fig.width=5, fig.height=6, out.width="4in", fig.cap="Example environmental time series.">>=
types <- c("sine", "arma", "regime", "linear", "constant")
x <- list()
for(i in 1:5) x[[i]] <- generate_env_ts(n_t = 100, type = types[i])
op <- par(mfrow = c(5, 1), mar = c(3,3,1,0), cex = 0.7)
for(i in 1:5) plot(x[[i]], type = "o", main = types[i])
par(op)
@

\section{Extra to be incorporated above}

We can visualize the variability in the Ricker $a$ parameters using the function \texttt{plot\_rickers} (Figure~\ref{fig:plot-rickers}).

<<plot-rickers, fig.width=7, fig.height=4.5, out.width="5in", fig.cap="Ricker curves from a sample of 40 years in the example simulation. Each panel represents a different stream population. Population 1 is cool tolerant and population 10 is warm tolerant. The colour of the Ricker curves represents the relative temperatue in that year (warm: red; cool: blue). The grey shaded area represents the variation in spawners observed within the simulation.">>=
pal <- rep("black", 10) # colours for the population IDs
plot_rickers(base1, pal = pal)
@

We can also use \texttt{metafolio} to sample from possible
investment weights (Figure~\ref{fig:monte-carlo-plot}). This type of
stochastic sampling is sometimes referred to as Monte Carlo
sampling.

<<monte-carlo-eg, cache=TRUE, results="hide">>=
set.seed(1)
weights_matrix <- create_asset_weights(n_pop = 5, n_sims = 500, 
  weight_lower_limit = 0.001)
mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
  n_sims = 500, mean_b = 1000)
@

<<monte-carlo-plot, fig.width=7, fig.height=4, out.width="5in", fig.cap="Efficient frontier of metapopulation portfolios (red dots). Each dot represents a different set of weights of the Ricker $b$ parameters. The colours in the right panel correspond to the five subpopulations with warm tolerant populations in warmer colours and cool tolerant populations in cooler colours.">>=
col_pal <- rev(gg_color_hue(5))
ef_dat <- plot_efficient_portfolios(port_vals = mc_ports$port_vals,
  pal = col_pal, weights_matrix = weights_matrix)
@

We can look at the correlation between salmon returns in the various
streams using the function
\texttt{plot\_correlation\_between\_returns}
(Figure~\ref{fig:return-correlations}).

<<return-correlations, fig.width=5, fig.height=5, out.width="5in", fig.cap="A plot comparing the log(returns) between each population. The subpopulation IDs are coloured from warm tolerant (warm colours) to cool tolerant (cool colours). Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics. Populations with thermal tolerance curves in the middle (e.g. population 6) are less correlated with other populations. Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity.">>=
plot_correlation_between_returns(base1)
@

\end{document}