### 2013-10-26 vignette/metafolio.Rnw
Make many edits to vignette

%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{metafolio}

% tw=68 for code
\documentclass[12pt]{article}
\usepackage{geometry}
\usepackage{url}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}

\setlength{\parskip}{12pt}
\setlength{\parindent}{0cm}

\begin{document}

<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
# set global chunk options
opts_chunk$set(fig.path='figure/', fig.align='center', fig.show='hold')
options(replace.assign=TRUE, width=90)
opts_knit$set(out.format = "latex")
opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, tidy=FALSE,
refresh=TRUE, cache=TRUE, autodep=TRUE)
#knit_theme$set("print")
@

\title{An introduction to the \texttt{metafolio} \texttt{R} package}
\author{Sean C. Anderson, and co-authors}
\date{}
\maketitle

The \texttt{metafolio} \texttt{R} package is a tool to simulate
metapopulations and apply financial portfolio optimization concepts to those
metapopulations. The package is primarily geared towards salmon, but could
be adapted for other taxonomic groups. This vignette accompanies the paper
\textit{Portfolio conservation of metapopulations under climate change}. In
this vignette we will describe the main functions in the \texttt{metafolio}
package, show how to re-create the main analyses in our paper, and demonstrate
some of the included plotting functions for exploring the output.

\section{Installing the package}

To install and load the package:

<<my-load-package, include=FALSE>>=
library(metafolio)
set.seed(1)
@

<<load-package, eval=FALSE>>=
install.packages("devtools") # if needed
devtools::install_github("metafolio", "seananderson")
library(metafolio)
@

You can get more detailed installation instructions, read the source code, and
report bugs on GitHub: \url{https://github.com/seananderson/metafolio}

\section{An example simulation}

The main simulation function is \texttt{meta\_sim}. We'll start by running a
simulation using the base-case parameter values from the paper. First, we'll
set up a list object that contains the argument values for the stationary
environmental stochasticity scenario. The arguments \texttt{mean\_value},
\texttt{ar}, and \texttt{ma} refer to the mean, autoregressive (AR), and
moving-average (MA) components of an ARIMA model. See \texttt{?arima} for
further explanation of ARIMA models in R. The argument \texttt{sigma\_env}
refers to the standard deviation of the environmental signal.

<<setup-arma>>=
arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2,
  ma = 0)
@

Next we'll run the simulation for one iteration. We'll simulate ten
populations and re-assess the fishery every five years.  Re-assessing the
fishery means that the simulation fits a Ricker curve to the observed
spawner-return data and updates the harvest-level targets.

<<base-case-run, cache=TRUE>>=
base1 <- meta_sim(n_pop = 10, env_params = arma_env_params,
  env_type = "arma", assess_freq = 5)
@

We can plot the output with the function \texttt{plot\_sim\_ts}. The output is
shown in Figure~\ref{fig:plot-base-case-ts}.

<<plot-base-case-ts, fig.width=5, fig.height=6.5, out.width="4in", fig.cap="An example simulation with stationary environmental stochasticity and the base-case parameter values.">>=
plot_sim_ts(base1, years_to_show = 70, burn = 30)
@

\clearpage

\section{Exploring prioritization strategies}

One of the key elements to the analysis in our paper is the choice of
which populations to prioritize for conservation. We can try different
prioritization strategies by manipulating the ``investment weights'' in each
stream of salmon. In the case of salmon, we represent these with the Ricker
$b_i$ parameters, which represent the maximum population capacity of streams
$i$ 1 through $n$.

In this example, we'll create a scenario in which we conserve response
diversity from across the spectrum of possible responses and a scenario
in which we conserve one half of the response diversity. We've set up the
weights carefully so that each metapopulation contains the same number of
populations (10) and the same total capacity. We set the Ricker $b_i$ parameters
equal to the nominal level of five salmon for the streams that we choose not to
prioritize.

<<spatial-plans>>=
w_plans <- list()
w_plans[["balanced"]] <- c(5, 1000, 5, 1000, 5, 5, 1000, 5,
    1000, 5)
w_plans[["one_half"]] <- c(rep(1000, 4), rep(5, 6))
w <- list()
for(i in 1:2) { # loop over plans
 w[[i]] <- list()
 for(j in 1:80) { # loop over iterations
   w[[i]][[j]] <- matrix(w_plans[[i]], nrow = 1)
 }
}
@

We've now created a nested list of stream capacities (\texttt{w}). The
first level of the list contains the two different plans. The second level
of the list contains the $b$ values for each iteration. Each iteration
will have re-sampled process noise and observation error when run with
\texttt{meta\_sim}. Here, we're keeping the $b$ values the same between
iterations, but that might not be the case if we wanted to stochastically
simulate the $b$ values. We're only going to run 80 iterations to reduce the
runtime of the example, but in reality you would likely want to run many more
iterations.

We can now stochastically simulate with these strategies using the function\\
\texttt{run\_cons\_plans} and plot the output with \texttt{plot\_cons\_plans}
(Figure~\ref{fig:spatial-runs}).

<<spatial-runs, fig.width=5, fig.height=5, out.width="4in", cache=TRUE, fig.cap="Two spatial conservation strategies shown in risk-return space with stationary environmental stochasticity. The dots show simulated metapopulations and the contours show 25\\% and 75\\% quantiles across 80 simulations per strategy. The grey line indicates the efficient frontier across all simulated metapopulations. The efficient frontier represents the minimum expected mean growth rate for a given expected variance in growth rate.">>=
set.seed(1)
arma_sp <- run_cons_plans(w, env_type = "arma", env_params =
  arma_env_params)
plot_cons_plans(arma_sp$plans_mv,
  plans_name = c("Balanced", "One half"),
  cols = c("#E41A1C", "#377EB8"), xlab = "Variance of growth rate",
  ylab = "Mean growth rate")
@

\section{Generating alternative environmental time series}

We can use the function \texttt{generate\_env\_ts} to create alternative
environmental time series. The function can generate five kinds of time
series: sine waves, regime shifts, linear changes, and constant values. Each
type has its own set of parameter arguments that are passed in a list
format. See \texttt{?generate\_env\_ts} for examples of all of these.

We can see demonstrations of all environmental time series types with the
default argument values with the following code (Figure~\ref{fig:env-eg}).

<<env-eg, fig.width=5, fig.height=6, out.width="4in", fig.cap="Example environmental time series.">>=
types <- c("sine", "arma", "regime", "linear", "constant")
x <- list()
for(i in 1:5) x[[i]] <- generate_env_ts(n_t = 100, type = types[i])
par(mfrow = c(5, 1), mar = c(3,3,1,0), cex = 0.7)
for(i in 1:5) plot(x[[i]], type = "o", main = types[i])
@

\clearpage

\section{Included plot types}

We can visualize the variability in the Ricker $a$ parameters using the
function \texttt{plot\_rickers} (Figure~\ref{fig:plot-rickers}).

<<plot-rickers, fig.width=7, fig.height=4.5, out.width="5in", fig.cap="Ricker curves from a sample of 40 years in the example simulation. Each panel represents a different stream population. Population 1 is more productive in cool conditions and population 10 is more producitive in warm conditions. The colour of the Ricker curves represents the relative temperatue in that year (warm: red; cool: blue). The grey shaded area represents the variation in spawners observed within the simulation.">>=
plot_rickers(base1, pal = rep("black", 10))
@

We can look at the correlation between salmon returns in the various
streams using the function \texttt{plot\_correlation\_between\_returns}
(Figure~\ref{fig:return-correlations}).

<<return-correlations, fig.width=5, fig.height=5, out.width="5in", fig.cap="A plot comparing the log(returns) between each population. The population IDs are coloured from warm tolerant (warm colours) to cool tolerant (cool colours). Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics. Populations with thermal tolerance curves in the middle (e.g.~population 6) are less correlated with other populations. Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity.">>=
plot_correlation_between_returns(base1)

@

\section{Optimizing metapopulation portfolios}

A standard procedure in financial portfolio management is to determine optimal
weights of the assets in a portfolio. The optimal portfolio weights are
referred to as the efficient frontier. This efficient frontier describes a
set of portfolios which have minimal risk for a specified level of return or
maximum return for a specified level of risk. While it would be complicated
to determine the optimal metapopulation portfolios via algebra we can do
so by letting \texttt{metafolio} sample from possible investment weights
(Figure~\ref{fig:monte-carlo-plot}) --- a form of Monte Carlo sampling.

<<monte-carlo-eg, eval=FALSE>>=
set.seed(999)
weights_matrix <- create_asset_weights(n_pop = 5, n_sims = 1000,
  weight_lower_limit = 0.001)
mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
  n_sims = 1000, mean_b = 1000)
@

<<monte-carlo-eg-load, eval=TRUE, echo=FALSE>>=
# To make the vignette compile more quickly:
weights_matrix <- create_asset_weights(n_pop = 5, n_sims = 1000,
  weight_lower_limit = 0.001)
load("port_vals.rda")
mc_ports <- list()
mc_ports$port_vals <- port_vals
@

<<monte-carlo-plot, fig.width=7, fig.height=4, out.width="5in", fig.cap="Efficient frontier of metapopulation portfolios (red dots). Each dot represents a different set of weights of the Ricker $b$ parameters. The colours in the right panel correspond to the five populations with warm tolerant populations in warmer colours and cool tolerant populations in cooler colours.">>=
col_pal <- rev(gg_color_hue(5))
ef_dat <- plot_efficient_portfolios(port_vals = mc_ports$port_vals,
  pal = col_pal, weights_matrix = weights_matrix)
@

\end{document}