### 2013-12-03 inst/ms/salmonfolios.tex
Move Fig and Table S code to som.md

\documentclass[11pt]{article}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[round]{natbib}
\bibpunct{(}{)}{;}{a}{}{;}
\textheight 20.5cm
%\topmargin -1.43cm
%\oddsidemargin 1.2cm
%\evensidemargin 1.2cm
%\textwidth 14.5cm
\usepackage{lineno}
\usepackage{xcolor}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\usepackage{booktabs}

\widowpenalty=1000
\clubpenalty=1000

\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}

\usepackage{fontspec}
\defaultfontfeatures{Ligatures=TeX}
%\setsansfont{Calibri}
%\setmonofont{Inconsolata}
%\usepackage{mathptmx}
%\setmainfont{Minion Pro}

% remove numbers in front of sections:
\makeatletter
\renewcommand\@seccntformat[1]{}
\makeatother

\hyphenation{meta-pop-ulation meta-pop-ulations sub-pop-ulations sub-pop-ulation e-con-o-mist en-vi-ron-men-tal pri-or-i-ti-za-tion}

\begin{document}

\linenumbers
\modulolinenumbers[2]
\begin{spacing}{1.4}
\setlength{\parskip}{12pt}
\setlength{\parindent}{0cm}
\include{ms}

\end{spacing}

\setlength{\parskip}{8pt}
\setlength{\parindent}{0cm}

\begin{spacing}{1.2}

\include{som}

\end{spacing}

\end{document}