### 2014-01-08 inst/ms/salmonfolios.tex
A variety of changes prior to prsb submission

\documentclass[11pt]{article}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{setspace}
%\usepackage[round]{natbib}
%\bibpunct{(}{)}{;}{a}{}{;}
\textheight 20.0cm
\usepackage{lineno}
\usepackage{xcolor}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\usepackage{booktabs}
\usepackage{cite}

\DeclareGraphicsExtensions{.pdf}

\widowpenalty=10000
\clubpenalty=10000

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}

% remove numbers in front of sections:
\makeatletter
\renewcommand\@seccntformat[1]{}
\makeatother

\hyphenation{meta-pop-ulation meta-pop-ulations sub-pop-ulations sub-pop-ulation e-con-o-mist en-vi-ron-men-tal pri-or-i-ti-za-tion}

\begin{document}
%\raggedright

\linenumbers
\modulolinenumbers[2]
\begin{spacing}{1.6}
\setlength{\parindent}{0cm}
\input{ms}

\bibliographystyle{prsb}
\bibliography{jshort,ms}

\clearpage

\section{Figure legends}

%\begin{center}
%\includegraphics[height=5.5in]{../examples/simulation-diagram3}
%\end{center}
\textbf{Fig.\ 1} Flow chart of the salmon-metapopulation simulation. There are $n$ salmon populations and $t$ generations. Blue text indicates values that are generated before the simulation progresses through time. Red text indicates steps in which calculations are performed through time. Black text indicates values that are calculated. Grey text indicates parameters that can be set. Green text indicates the looping structure of the simulation.

%\clearpage

%\begin{center}
%\includegraphics[width=2.9in]{../examples/thermal-curve-scenarios}
%\end{center}
\textbf{Figure 2.} Different ways of prioritizing response-diversity conservation. Panel a shows thermal tolerance curves for ten possible populations and panels b--e show different ways of prioritizing four of those populations. The curves describe how productivity varies with temperature for a given population. Some populations thrive at low temperatures (cool colours) and some at warm temperatures (warm colours). Some are tolerant to a wider range of environmental conditions (yellow-to-green colours) but with a lower maximum productivity. The total possible productivity (the area under the curves) is the same for each population.

%\clearpage

%\begin{center}
%\includegraphics[width=4.0in]{../examples/spatial-arma-sim-full}
%\end{center}
\textbf{Figure 3.} The components of an example metapopulation simulation. We show, from top to bottom, the temperature signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve. The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) temperatures.

%\clearpage

%\begin{center}
%\includegraphics[width=4.5in]{../examples/spatial-mv}
%\end{center}
\textbf{Figure 4.} The importance of preserving environmental response diversity through spatial conservation strategies. The conservation strategies correspond to Figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink). In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations. The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy. We also show example metapopulation abundance time series for the (c, e) short-term and (d, f) long-term environmental-fluctuation scenarios. The grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.

%\clearpage

%\begin{center}
%\includegraphics[width=4.5in]{../examples/cons-plans-n}
%\end{center}
\textbf{Figure 5.} The importance of preserving as many populations as possible when we don't know how response diversity is distributed. In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations. We show metapopulations in which 2 (red), 4 (orange), 8 (yellow), or 16 (green) populations of random response diversity are conserved. The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy. We also show example metapopulation (c) rate-of-change and (d) abundance time series for the short-term environmental-fluctuation scenario. The grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate

\clearpage

\end{spacing}

\setlength{\parskip}{8pt}
\setlength{\parindent}{0cm}

\begin{spacing}{1.0}

\include{som}

\end{spacing}

\end{document}