### 2014-07-10 inst/ms/appendices/Appendix-A.tex
Add separate appendix files

\documentclass[12pt]{article}

\setlength\parindent{0in}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. Ecological Applications. VOL:pp--pp.}

\bigskip

\textsc{Appendix A.} The \texttt{metafolio} \textsf{R} package.

\bigskip

The \texttt{metafolio} \textsf{R} package contains the functions and code to carry out the analyses in our paper. \texttt{metafolio}. The package can be installed from CRAN with:

\begin{verbatim}
install.packages("metafolio")
\end{verbatim}

%Alternatively, you can view the code and install the package from\\ \texttt{http://github.com/seananderson/metafolio}
%\bigskip

The included vignette describes the package and illustrates some example simulations. You can view the vignette with:

\begin{verbatim}
vignette("metafolio")
\end{verbatim}

You can view the help for the package with:

\begin{verbatim}
?metafolio
help(package = "metafolio")
\end{verbatim}

The figures from this paper can be re-created by downloading the code from GitHub and sourcing the file \texttt{README.R} in the \texttt{inst/examples} folder:

\begin{verbatim}
setwd("metafolio/inst/examples")
source("README.R")
\end{verbatim}

\end{document}