### 2014-07-10 inst/ms/appendices/Appendix-B.tex

\documentclass[12pt]{article}

\usepackage[round]{natbib}
\bibpunct{}{}{;}{a}{}{}
\setlength\parindent{0in}
\usepackage[originalparameters]{ragged2e}
\usepackage{array}
\usepackage{booktabs}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. Ecological Applications. VOL:pp--pp.}

\bigskip

\textsc{Appendix B.} Simulation input parameters and default values.

\bigskip

\begin{table}[h!]
\centering
\footnotesize
\caption{Input parameters to the salmon metapopulation simulation with default values.}
\smallskip
\begin{tabular}{>{\RaggedRight}p{7.8cm}p{1.1cm}p{2.5cm}>{\RaggedRight}p{3.3cm}}
\toprule
Description                                                          & Symbol                & Value                  & Reference  \\
\midrule

\textit{Population dynamics parameters}                              &                       &                        &             \\
Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.7                    & \citep{thorson2014a}  \\
AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.4                    & \citep{thorson2014a}  \\
Fraction of fish that stray from natal streams                       & $f_{\mathrm{stray}}$  & 0.02                   & \citep{quinn2005} and references therein  \\
Exponential rate of decay of straying with distance                  & $m$                   & 0.1                    & \citep{cooper1999}      \\
Range of maximum productivities                                      & $a_i^{\mathrm{max}}$  & 2.2--2.9             &  \citep{dorner2008}   \\

\noalign{\vskip 3mm}
\textit{Environmental parameters}                                    &                       &                        &             \\
Width parameter for thermal-tolerance curves for populations $i$ 1 to $n$  (values generate widths in line with listed references) & $W_i$                 & 0.08--0.04--0.08       &  \citep{brett1952, eliason2011}           \\
Optimum environmental value for populations $i$ 1 to $n$             & $e_i^{\mathrm{opt}}$  & 13--19                 &  \citep{eliason2011}           \\
Standard deviation of annual temperature fluctuations          & $\sigma_d$            & 2                      & \citep{eliason2011}      \\
AR(1) autocorrelation of annual temperature fluctuations       & $\rho_e$              & 0.1                    &             \\
Annual increase in stream temperature in degrees Celcius             & $\beta_e$             & 0.04                   & \citep{mantua2010}     \\

\noalign{\vskip 3mm}
\textit{Fishery parameters}                                          &                       &                        &             \\
Standard deviation of beta distribution for implementation error     & $\sigma_{h}$          & 0.1                    & \citep{pestes2008}            \\
Frequency of assessment (years)                                      & $f_{\mathrm{assess}}$ & 5                      &             \\
\bottomrule
\end{tabular}
\label{t:pars}
\end{table}
\clearpage

%\bibliographystyle{ecology3}
%\bibliography{jshort,ms}
\begin{thebibliography}{8}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{{\tt #1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi

\bibitem[{Brett(1952)}]{brett1952}
Brett, J.~R.
\newblock 1952.
\newblock Temperature Tolerance in Young Pacific Salmon, Genus
  \textit{Oncorhynchus}.
\newblock J. Fish. Res. Bd. Can. 9:265--323.

\bibitem[{Cooper and Mangel(1999)}]{cooper1999}
Cooper, A.~B., and M.~Mangel.
\newblock 1999.
\newblock The dangers of ignoring metapopulation structure for the conservation
  of salmonids.
\newblock Fish. Bull. 97:213--226.

\bibitem[{Dorner et~al.(2008)Dorner, Peterman, and Haeseker}]{dorner2008}
Dorner, B., R.~M. Peterman, and S.~L. Haeseker.
\newblock 2008.
\newblock Historical trends in productivity of 120 {Pacific} pink, chum, and
  sockeye salmon stocks reconstructed by using a {Kalman} filter.
\newblock Can. J. Fish. Aquat. Sci. 65:1842--1866.

\bibitem[{Eliason et~al.(2011)Eliason, Clark, Hague, Hanson, Gallagher,
  Jeffries, Gale, Patterson, Hinch, and Farrell}]{eliason2011}
Eliason, E.~J., T.~D. Clark, M.~J. Hague, L.~M. Hanson, Z.~S. Gallagher, K.~M.
  Jeffries, M.~K. Gale, D.~A. Patterson, S.~G. Hinch, and A.~P. Farrell.
\newblock 2011.
\newblock Differences in thermal tolerance among sockeye salmon populations.
\newblock Science 332:109--112.

\bibitem[{Mantua et~al.(2010)Mantua, Tohver, and Hamlet}]{mantua2010}
Mantua, N., I.~Tohver, and A.~Hamlet.
\newblock 2010.
\newblock Climate change impacts on streamflow extremes and summertime stream
  temperature and their possible consequences for freshwater salmon habitat in
  {Washington State}.
\newblock Climatic Change 102:187--223.

\bibitem[{Pestes et~al.(2008)Pestes, Peterman, Bradford, and Wood}]{pestes2008}
Pestes, L.~R., R.~M. Peterman, M.~J. Bradford, and C.~C. Wood.
\newblock 2008.
\newblock Bayesian decision analysis for evaluating management options to
  promote recovery of a depleted salmon population.
\newblock Conserv. Biol. 22:351--361.

\bibitem[{Quinn(2005)}]{quinn2005}
Quinn, T.~P.
\newblock 2005.
\newblock The Behaviour and Ecology of Pacific Salmon and Trout.
\newblock American Fisheries Society, Bethesda, Maryland.

\bibitem[{Thorson et~al.(2014)Thorson, Jensen, and Zipkin}]{thorson2014a}
Thorson, J.~T., O.~P. Jensen, and E.~F. Zipkin.
\newblock 2014.
\newblock How variable is recruitment for exploited marine fishes? A
  hierarchical model for testing life history theory.
\newblock Can. J. Fish. Aquat. Sci. 71:973--983.

\end{thebibliography}

\end{document}