### 2014-07-10 inst/ms/appendices/Appendix-C.tex

\documentclass[12pt]{article}
\usepackage{graphicx}

\setlength\parindent{0in}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. Ecological Applications. VOL:pp--pp.}

\bigskip

\textsc{Appendix C.} An example straying matrix.

\bigskip

\includegraphics[width=3.5in]{Fig-C1}\\
\textsc{Fig.~C1.} An example straying matrix. The rows and columns represent different populations (indicated by population number). Dark blue indicates a high rate of straying and light blue indicates a low rate of straying.

\end{document}