### 2014-07-10 inst/ms/appendices/Appendix-F.tex

\documentclass[12pt]{article}
\usepackage{graphicx}

\setlength\parindent{0in}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. Ecological Applications. VOL:pp--pp.}

\bigskip

\textsc{Appendix F.} Example simulated time series from alternative conservation scenarios.

\bigskip

\clearpage

\includegraphics[width=4.3in]{Fig-F1}\\
\textsc{Fig.~F1.} Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations. This is the same as Fig.~3 but in colour.
\clearpage

\includegraphics[width=4.3in]{Fig-F2}\\
\textsc{Fig.~F2.} Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations.
\clearpage

\includegraphics[width=4.3in]{Fig-F3}\\
\textsc{Fig.~F3.} Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-F4}\\
\textsc{Fig.~F4.} Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-F5}\\
\textsc{Fig.~F5.} \textbf{Two populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.
\clearpage

\includegraphics[width=4.3in]{Fig-F6}\\
\textsc{Fig.~F6.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.
\clearpage

\includegraphics[width=4.3in]{Fig-F7}\\
\textsc{Fig.~F7.} \textbf{Two populations} conserved with random response diversity and \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-F8}\\
\textsc{Fig.~F8.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-F9}\\
\textsc{Fig.~F9.} \textbf{Two populations} conserved with random response diversity and \textbf{long-term declining stream flow}.
\clearpage

\includegraphics[width=4.3in]{Fig-F10}\\
\textsc{Fig.~F10.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term declining stream flow}.

\end{document}