### 2014-07-10 inst/ms/anderson-etal-metafolio.tex
Get main .tex file ready to submit

\documentclass[12pt,english]{article}
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
\geometry{letterpaper}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[round]{natbib}
\bibpunct{(}{)}{;}{a}{}{;}
\textheight 22.0cm
\usepackage{lineno}
\usepackage{xcolor}
%\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\usepackage{booktabs}
\usepackage{cite}
\usepackage[originalparameters]{ragged2e}
\usepackage{array}
%\usepackage{mathptmx}

\newlabel{f:flowchart}{{1}{999}}
\newlabel{f:curves}{{2}{999}}
\newlabel{f:ts}{{3}{999}}
\newlabel{f:sp}{{4}{999}}
\newlabel{f:n}{{5}{999}}
\newlabel{f:squeeze}{{6}{999}}

\newcommand{\somR}{Appendix A}
\newcommand{\somparam}{Appendix B}
\newcommand{\somstray}{Appendix C}
\newcommand{\somsens}{Appendix D}
\newcommand{\somcor}{Appendix E}
\newcommand{\somts}{Appendix F}

% SOM:
\newlabel{t:pars}{{B1}{2}}
\newlabel{f:stray}{{C1}{3}}
\newlabel{f:eg-sens}{{D1}{4}}
\newlabel{f:ret-corr}{{E1}{5}}
\newlabel{f:eg-sp-arma-full}{{F1}{6}}
\newlabel{f:eg-sp-arma-half}{{F2}{7}}
\newlabel{f:eg-sp-linear-full}{{F3}{8}}
\newlabel{f:eg-sp-linear-half}{{F4}{9}}
\newlabel{f:eg-n-arma-two}{{F5}{10}}
\newlabel{f:eg-n-arma-sixteen}{{F6}{11}}
\newlabel{f:eg-n-linear-two}{{F7}{12}}
\newlabel{f:eg-n-linear-sixteen}{{F8}{13}}
\newlabel{f:eg-n-squeeze-two}{{F9}{14}}
\newlabel{f:eg-n-squeeze-twelve}{{F10}{15}}

\DeclareGraphicsExtensions{.pdf}

\widowpenalty=10000
\clubpenalty=10000

%\usepackage{titlesec}
%\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
%\titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}

\title{Portfolio conservation of metapopulations under\\climate change}

\author{
Sean C. Anderson$^{1\ast}$ \and Jonathan W. Moore$^{1,2}$ \and Michelle M. McClure$^3$ \and
Nicholas K. Dulvy$^1$ \and Andrew B. Cooper$^2$
}
\date{}

% remove numbers in front of sections:
\makeatletter
\renewcommand\@seccntformat[1]{}
\makeatother

\hyphenation{meta-pop-ulation meta-pop-ulations sub-pop-ulations sub-pop-ulation e-con-o-mist en-vi-ron-men-tal pri-or-i-ti-za-tion}

%\newcommand{\R}[1]{\label{#1}\linelabel{#1}}
%\newcommand{\lr}[1]{line~\lineref{#1}}

\begin{document}
\raggedright

\begin{spacing}{1.1}
\setlength\parskip{0.11in}
\setlength\parindent{0in}
%\include{response}
\setlength\parskip{0in}
\end{spacing}

\setcounter{page}{1}
\linenumbers

%\modulolinenumbers[2]
\begin{spacing}{1.9}
%\begin{spacing}{1.2}
\setlength{\parindent}{0.9cm}

\maketitle

\input{ms}

%\bibliographystyle{ecology3}
%\bibliography{jshort,ms}
%\input{anderson-etal-metafolio.bbl}
\begin{thebibliography}{64}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{{\tt #1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi

\bibitem[{Allendorf et~al.(1997)Allendorf, Bayles, Bottom, Currens, Frissell,
  Hankin, Lichatowich, Nehlsen, Trotter, and Williams}]{allendorf1997}
Allendorf, F.~W., D.~Bayles, D.~L. Bottom, K.~P. Currens, C.~A. Frissell,
  D.~Hankin, J.~A. Lichatowich, W.~Nehlsen, P.~C. Trotter, and T.~H. Williams.
\newblock 1997.
\newblock Prioritizing {Pacific} salmon stocks for conservation.
\newblock Conserv. Biol. 11:140--152.

\bibitem[{Anderson(2014)}]{metafoliopkg}
Anderson, S.~C. 2014.
\newblock metafolio: Salmon metapopulation simulations for conservation
  planning through portfolio optimization. R package version 0.1.0.
  http://cran.r-project.org/package=metafolio.

\bibitem[{Anderson et~al.(2013)Anderson, Cooper, and Dulvy}]{anderson2013}
Anderson, S.~C., A.~B. Cooper, and N.~K. Dulvy.
\newblock 2013.
\newblock Ecological prophets: Quantifying metapopulation portfolio effects.
\newblock Methods Ecol. Evol. 4:971--981.

\bibitem[{Ando and Mallory(2012)}]{ando2012}
Ando, A.~W., and M.~L. Mallory.
\newblock 2012.
\newblock Optimal portfolio design to reduce climate-related conservation
  uncertainty in the {Prairie Pothole Region}.
\newblock Proc. Natl. Acad. Sci. U.S.A. 109:6484--6489.

\bibitem[{Balvanera et~al.(2006)Balvanera, Pfisterer, Buchmann, He,
  Nakashizuka, Raffaelli, and Schmid}]{balvanera2006}
Balvanera, P., A.~B. Pfisterer, N.~Buchmann, J.-S. He, T.~Nakashizuka,
  D.~Raffaelli, and B.~Schmid.
\newblock 2006.
\newblock Quantifying the evidence for biodiversity effects on ecosystem
  functioning and services.
\newblock Ecol. Lett. 9:1146--1156.

\bibitem[{Battin et~al.(2007)Battin, Wiley, Ruckelshaus, Palmer, Korb, Bartz,
  and Imaki}]{battin2007}
Battin, J., M.~W. Wiley, M.~H. Ruckelshaus, R.~N. Palmer, E.~Korb, K.~K. Bartz,
  and H.~Imaki.
\newblock 2007.
\newblock Projected impacts of climate change on salmon habitat restoration.
\newblock Proc. Natl. Acad. Sci. U.S.A. 104:6720--6725.

\bibitem[{Beechie et~al.(2013)Beechie, Imaki, Greene, Wade, Wu, Pess, Roni,
  Kimball, Stanford, Kiffney, and Mantua}]{beechie2013}
Beechie, T., H.~Imaki, J.~Greene, A.~Wade, H.~Wu, G.~Pess, P.~Roni, J.~Kimball,
  J.~Stanford, P.~Kiffney, and N.~Mantua.
\newblock 2013.
\newblock Restoring salmon habitat for a changing climate.
\newblock River Res. Applic. 29:939--960.

\bibitem[{Carlson and Satterthwaite(2011)}]{carlson2011}
Carlson, S.~M., and W.~H. Satterthwaite.
\newblock 2011.
\newblock Weakened portfolio effect in a collapsed salmon population complex.
\newblock Can. J. Fish. Aquat. Sci. 68:1579--1589.

\bibitem[{Cooper and Mangel(1999)}]{cooper1999}
Cooper, A.~B., and M.~Mangel.
\newblock 1999.
\newblock The dangers of ignoring metapopulation structure for the conservation
  of salmonids.
\newblock Fish. Bull. 97:213--226.

\bibitem[{Crozier et~al.(2008)Crozier, Hendry, Lawson, Quinn, Mantua, Battin,
  Shaw, and Huey}]{crozier2008}
Crozier, L.~G., A.~P. Hendry, P.~W. Lawson, T.~Quinn, N.~J. Mantua, J.~Battin,
  R.~G. Shaw, and R.~B. Huey.
\newblock 2008.
\newblock Potential responses to climate change in organisms with complex life
  histories: evolution and plasticity in {Pacific} salmon.
\newblock Evol. Appl. 1:252--270.

\bibitem[{de~Mazancourt et~al.(2013)de~Mazancourt, Isbell, Larocque, Berendse,
  De~Luca, Grace, Haegeman, Wayne~Polley, Roscher, Schmid, Tilman, van Ruijven,
  Weigelt, Wilsey, and Loreau}]{demazancourt2013}
de~Mazancourt, C., F.~Isbell, A.~Larocque, F.~Berendse, E.~De~Luca, J.~B.
  Grace, B.~Haegeman, H.~Wayne~Polley, C.~Roscher, B.~Schmid, D.~Tilman, J.~van
  Ruijven, A.~Weigelt, B.~J. Wilsey, and M.~Loreau.
\newblock 2013.
\newblock Predicting ecosystem stability from community composition and
  biodiversity.
\newblock Ecol. Lett. 16:617--625.

\bibitem[{DFO(2005)}]{dfo2005wsp}
DFO. 2005.
\newblock Canada's policy for conservation of wild {Pacific} salmon.
\newblock Technical report, Fisheries and Oceans Canada, Vancouver, BC.

\bibitem[{Dorner et~al.(2008)Dorner, Peterman, and Haeseker}]{dorner2008}
Dorner, B., R.~M. Peterman, and S.~L. Haeseker.
\newblock 2008.
\newblock Historical trends in productivity of 120 {Pacific} pink, chum, and
  sockeye salmon stocks reconstructed by using a {Kalman} filter.
\newblock Can. J. Fish. Aquat. Sci. 65:1842--1866.

\bibitem[{Eliason et~al.(2011)Eliason, Clark, Hague, Hanson, Gallagher,
  Jeffries, Gale, Patterson, Hinch, and Farrell}]{eliason2011}
Eliason, E.~J., T.~D. Clark, M.~J. Hague, L.~M. Hanson, Z.~S. Gallagher, K.~M.
  Jeffries, M.~K. Gale, D.~A. Patterson, S.~G. Hinch, and A.~P. Farrell.
\newblock 2011.
\newblock Differences in thermal tolerance among sockeye salmon populations.
\newblock Science 332:109--112.

\bibitem[{Elmqvist et~al.(2003)Elmqvist, Folke, Nystr{\"o}m, Peterson,
  Bengtsson, Walker, and Norberg}]{elmqvist2003}
Elmqvist, T., C.~Folke, M.~Nystr{\"o}m, G.~Peterson, J.~Bengtsson, B.~Walker,
  and J.~Norberg.
\newblock 2003.
\newblock Response diversity, ecosystem change, and resilience.
\newblock Front. Ecol. Environ. 1:488--494.

\bibitem[{Figge(2004)}]{figge2004}
Figge, F.
\newblock 2004.
\newblock Bio-folio: applying portfolio theory to biodiversity.
\newblock Biodivers. Conserv. 13:827--849.

\bibitem[{Fraser et~al.(2011)Fraser, Weir, Bernatchez, Hansen, and
  Taylor}]{fraser2011}
Fraser, D.~J., L.~K. Weir, L.~Bernatchez, M.~M. Hansen, and E.~B. Taylor.
\newblock 2011.
\newblock Extent and scale of local adaptation in salmonid fishes: review and
  meta-analysis.
\newblock Heredity 106:404--420.

\bibitem[{Gustafson et~al.(2007)Gustafson, Waples, Myers, Weitkamp, Bryant,
  Johnson, and Hard}]{gustafson2007}
Gustafson, R.~G., R.~S. Waples, J.~M. Myers, L.~A. Weitkamp, G.~J. Bryant,
  O.~W. Johnson, and J.~J. Hard.
\newblock 2007.
\newblock Pacific salmon extinctions: quantifying lost and remaining diversity.
\newblock Conserv. Biol. 21:1009--1020.

\bibitem[{Haak and Williams(2012)}]{haak2012}
Haak, A.~L., and J.~E. Williams.
\newblock 2012.
\newblock Spreading the risk: native trout management in a warmer and
  less-certain future.
\newblock N. Am. J. Fish. Manage. 32:387--401.

\bibitem[{Hannah et~al.(2002)Hannah, Midgley, Lovejoy, Bond, Bush, Lovett,
  Scott, and Woodward}]{hannah2002a}
Hannah, L., G.~F. Midgley, T.~Lovejoy, W.~J. Bond, M.~Bush, J.~C. Lovett,
  D.~Scott, and F.~I. Woodward.
\newblock 2002.
\newblock Conservation of biodiversity in a changing climate.
\newblock Conserv. Biol. 16:264--268.

\bibitem[{Hilborn et~al.(2003)Hilborn, Quinn, Schindler, and
  Rogers}]{hilborn2003}
Hilborn, R.~W., T.~Quinn, D.~Schindler, and D.~Rogers.
\newblock 2003.
\newblock Biocomplexity and fisheries sustainability.
\newblock Proc. Natl. Acad. Sci. U.S.A. 100:6564--6568.

\bibitem[{Hilborn and Walters(1992)}]{hilborn1992}
Hilborn, R.~W., and C.~Walters.
\newblock 1992.
\newblock Quantitative Fisheries Stock Assessment: Choice, Dynamics, and
  Uncertainty.
\newblock Chapman and Hall, London.

\bibitem[{Isaak and Rieman(2013)}]{isaak2013}
Isaak, D.~J., and B.~E. Rieman.
\newblock 2013.
\newblock Stream isotherm shifts from climate change and implications for
  distributions of ectothermic organisms.
\newblock Global Change Biol. 19:742--751.

\bibitem[{Isaak et~al.(2007)Isaak, Thurow, Rieman, and Dunham}]{isaak2007}
Isaak, D.~J., R.~F. Thurow, B.~E. Rieman, and J.~B. Dunham.
\newblock 2007.
\newblock Chinook salmon use of spawning patches: relative roles of habitat
  quality, size, and connectivity.
\newblock Ecol. Appl. 17:352--364.

\bibitem[{Isaak et~al.(2012)Isaak, Wollrab, Horan, and Chandler}]{isaak2012}
Isaak, D.~J., S.~Wollrab, D.~Horan, and G.~Chandler.
\newblock 2012.
\newblock Climate change effects on stream and river temperatures across the
  northwest U.S. from 1980--2009 and implications for salmonid fishes.
\newblock Climatic Change 113:499--524.

\bibitem[{Ives and Carpenter(2007)}]{ives2007}
Ives, A.~R., and S.~R. Carpenter.
\newblock 2007.
\newblock Stability and diversity of ecosystems.
\newblock Science 317:58--62.

\bibitem[{Jentsch et~al.(2007)Jentsch, Kreyling, and
  Beierkuhnlein}]{jentsch2007}
Jentsch, A., J.~Kreyling, and C.~Beierkuhnlein.
\newblock 2007.
\newblock A new generation of climate-change experiments: events, not trends.
\newblock Front. Ecol. Environ. 5:365--374.

\bibitem[{Johnson and Schindler(2013)}]{johnson2013a}
Johnson, S.~P., and D.~E. Schindler.
\newblock 2013.
\newblock Marine trophic diversity in an anadromous fish is linked to its
  life-history variation in fresh water.
\newblock Biol. Lett. 9:20120824.

\bibitem[{Koellner and Schmitz(2006)}]{koellner2006}
Koellner, T., and O.~J. Schmitz.
\newblock 2006.
\newblock Biodiversity, ecosystem function, and investment risk.
\newblock BioScience 56:977--985.

\bibitem[{Lackey(2003)}]{lackey2003}
Lackey, R.~T.
\newblock 2003.
\newblock {Pacific Northwest} salmon: forecasting their status in 2100.
\newblock Rev. Fish. Sci. 11:35--88.

\bibitem[{Lee and Jetz(2008)}]{lee2008}
Lee, T.~M., and W.~Jetz.
\newblock 2008.
\newblock Future battlegrounds for conservation under global change.
\newblock Proc. R. Soc. Lond. B 275:1261--1270.

\bibitem[{Luce et~al.(2013)Luce, Abatzoglou, and Holden}]{luce2013}
Luce, C.~H., J.~T. Abatzoglou, and Z.~A. Holden.
\newblock 2013.
\newblock The missing mountain water: slower westerlies decrease orographic
  enhancement in the {Pacific Northwest USA}.
\newblock Science 342:1360--1364.

\bibitem[{Luce and Holden(2009)}]{luce2009}
Luce, C.~H., and Z.~A. Holden.
\newblock 2009.
\newblock Declining annual streamflow distributions in the {Pacific Northwest
  United States}, 1948--2006.
\newblock Geophys. Res. Lett. 36.

\bibitem[{Mantua et~al.(2010)Mantua, Tohver, and Hamlet}]{mantua2010}
Mantua, N., I.~Tohver, and A.~Hamlet.
\newblock 2010.
\newblock Climate change impacts on streamflow extremes and summertime stream
  temperature and their possible consequences for freshwater salmon habitat in
  {Washington State}.
\newblock Climatic Change 102:187--223.

\bibitem[{Margules and Pressey(2000)}]{margules2000}
Margules, C.~R., and R.~L. Pressey.
\newblock 2000.
\newblock Systematic conservation planning.
\newblock Nature 405:243--253.

\bibitem[{Markowitz(1952)}]{markowitz1952}
Markowitz, H.
\newblock 1952.
\newblock Portfolio Selection.
\newblock J. Finance 7:77--91.

\bibitem[{McClure et~al.(2008{\natexlab{{\em a\/}}})McClure, Carlson, Beechie,
  Pess, Jorgensen, Sogard, Sultan, Holzer, Travis, Sanderson, Power, and
  Carmichael}]{mcclure2008a}
McClure, M.~M., S.~M. Carlson, T.~J. Beechie, G.~R. Pess, J.~C. Jorgensen,
  S.~M. Sogard, S.~E. Sultan, D.~M. Holzer, J.~Travis, B.~L. Sanderson, M.~E.
  Power, and R.~W. Carmichael.
\newblock 2008{\natexlab{{\em a\/}}}.
\newblock Evolutionary consequences of habitat loss for {Pacific} anadromous
  salmonids.
\newblock Evol. Appl. 1:300--318.

\bibitem[{McClure et~al.(2008{\natexlab{{\em b\/}}})McClure, Utter, Baldwin,
  Carmichael, Hassemer, Howell, Spruell, Cooney, Schaller, and
  Petrosky}]{mcclure2008b}
McClure, M.~M., F.~M. Utter, C.~Baldwin, R.~W. Carmichael, P.~F. Hassemer,
  P.~J. Howell, P.~Spruell, T.~D. Cooney, H.~A. Schaller, and C.~E. Petrosky.
\newblock 2008{\natexlab{{\em b\/}}}.
\newblock Evolutionary effects of alternative artificial propagation programs:
  implications for viability of endangered anadromous salmonids.
\newblock Evol. Appl. 1:356--375.

\bibitem[{McCullough(1999)}]{mccullough1999}
McCullough, D.~A., 1999.
\newblock A review and synthesis of effects of alterations to the water
  temperature regime on freshwater life stages of salmonids, with special
  reference to {Chinook} salmon.
\newblock EPA 910-R-99-010, US Environmental Protection Agency.

\bibitem[{McElhany et~al.(2000)McElhany, Ruckelshaus, Ford, Wainwright, and
  Bjorkstedt}]{mcelhany2000}
McElhany, P., M.~H. Ruckelshaus, M.~J. Ford, T.~C. Wainwright, and E.~P.
  Bjorkstedt, 2000.
\newblock {Viable Salmonid Populations} and the recovery of {Evolutionarily
  Significant Units}.
\newblock Technical Report NOAA Tech. Memo. NMFS-NWFSC-42, U.S. Deptartment of
  Commerce.

\bibitem[{Moore et~al.(2010)Moore, Mcclure, Rogers, and Schindler}]{moore2010}
Moore, J.~W., M.~Mcclure, L.~A. Rogers, and D.~E. Schindler.
\newblock 2010.
\newblock Synchronization and portfolio performance of threatened salmon.
\newblock Conserv. Lett. 3:340--348.

\bibitem[{Mori et~al.(2013)Mori, Furukawa, and Sasaki}]{mori2013}
Mori, A.~S., T.~Furukawa, and T.~Sasaki.
\newblock 2013.
\newblock Response diversity determines the resilience of ecosystems to
  environmental change.
\newblock Biol. Rev. 88:349--364.

\bibitem[{Patterson et~al.(2007)Patterson, Skibo, Barnes, Hills, and
  Macdonald}]{patterson2007}
Patterson, D.~A., K.~M. Skibo, D.~P. Barnes, J.~A. Hills, and J.~S. Macdonald.
\newblock 2007.
\newblock The influence of water temperature on time to surface for adult
  sockeye salmon carcasses and the limitations in estimating salmon carcasses
  in the {Fraser River, British Columbia}.
\newblock N. Am. J. Fish. Manage. 27:878--884.

\bibitem[{Pestes et~al.(2008)Pestes, Peterman, Bradford, and Wood}]{pestes2008}
Pestes, L.~R., R.~M. Peterman, M.~J. Bradford, and C.~C. Wood.
\newblock 2008.
\newblock Bayesian decision analysis for evaluating management options to
  promote recovery of a depleted salmon population.
\newblock Conserv. Biol. 22:351--361.

\bibitem[{Policansky and Magnuson(1998)}]{policansky1998}
Policansky, D., and J.~J. Magnuson.
\newblock 1998.
\newblock Genetics, metapopulations, and ecosystem management of fisheries.
\newblock Ecol. Appl. 8:S119--S123.

\bibitem[{Pressey et~al.(2007)Pressey, Cabeza, Watts, Cowling, and
  Wilson}]{pressey2007}
Pressey, R.~L., M.~Cabeza, M.~E. Watts, R.~M. Cowling, and K.~A. Wilson.
\newblock 2007.
\newblock Conservation planning in a changing world.
\newblock Trends Ecol. Evolut. 22:583--592.

\bibitem[{Quinn(2005)}]{quinn2005}
Quinn, T.~P.
\newblock 2005.
\newblock The Behaviour and Ecology of Pacific Salmon and Trout.
\newblock American Fisheries Society, Bethesda, Maryland.

\bibitem[{{\textsf{R} Core Team}(2013)}]{r2013}
{\textsf{R} Core Team}, 2013.
\newblock \textsf{R}: {A} language and environment for statistical computing.
\newblock \textsf{R} Foundation for Statistical Computing, Vienna, Austria.

\bibitem[{Rogers and Schindler(2008)}]{rogers2008}
Rogers, L.~A., and D.~E. Schindler.
\newblock 2008.
\newblock Asynchrony in population dynamics of sockeye salmon in southwest
  {Alaska}.
\newblock Oikos 117:1578--1586.

\bibitem[{Ruckelshaus et~al.(2002)Ruckelshaus, Levin, Johnson, and
  Kareiva}]{ruckelshaus2002}
Ruckelshaus, M.~H., P.~Levin, J.~B. Johnson, and P.~M. Kareiva.
\newblock 2002.
\newblock The {Pacific} salmon wars: what science brings to the challenge of
  recovering species.
\newblock Annu. Rev. Ecol. Syst. 33:665--706.

\bibitem[{Scheuerell et~al.(2006)Scheuerell, Hilborn, Ruckelshaus, Bartz,
  Lagueux, Haas, and Rawson}]{scheuerell2006}
Scheuerell, M.~D., R.~Hilborn, M.~H. Ruckelshaus, K.~K. Bartz, K.~M. Lagueux,
  A.~D. Haas, and K.~Rawson.
\newblock 2006.
\newblock The {Shiraz} model: a tool for incorporating anthropogenic effects
  and fish--habitat relationships in conservation planning.
\newblock Can. J. Fish. Aquat. Sci. 63:1596--1607.

\bibitem[{Schindler et~al.(2008)Schindler, Augerot, Fleishman, Mantua, Riddell,
  Ruckelshaus, Seeb, and Webster}]{schindler2008}
Schindler, D.~E., X.~Augerot, E.~Fleishman, N.~J. Mantua, B.~Riddell,
  M.~Ruckelshaus, J.~Seeb, and M.~Webster.
\newblock 2008.
\newblock Climate change, ecosystem impacts, and management for {Pacific}
  salmon.
\newblock Fisheries 33:502--506.

\bibitem[{Schindler et~al.(2010)Schindler, Hilborn, Chasco, Boatright, Quinn,
  Rogers, and Webster}]{schindler2010}
Schindler, D.~E., R.~Hilborn, B.~Chasco, C.~P. Boatright, T.~P. Quinn, L.~A.
  Rogers, and M.~S. Webster.
\newblock 2010.
\newblock Population diversity and the portfolio effect in an exploited
  species.
\newblock Nature 465:609--612.

\bibitem[{Schtickzelle and Quinn(2007)}]{schtickzelle2007}
Schtickzelle, N., and T.~P. Quinn.
\newblock 2007.
\newblock A metapopulation perspective for salmon and other anadromous fish.
\newblock Fish Fish. 8:297--314.

\bibitem[{Stockwell et~al.(2003)Stockwell, Hendry, and
  Kinnison}]{stockwell2003}
Stockwell, C., A.~Hendry, and M.~Kinnison.
\newblock 2003.
\newblock Contemporary evolution meets conservation biology.
\newblock Trends. Ecol. Evol. 18:94--101.

\bibitem[{Thibaut et~al.(2012)Thibaut, Connolly, and Sweatman}]{thibaut2012}
Thibaut, L., S.~R. Connolly, and H.~P.~A. Sweatman.
\newblock 2012.
\newblock Diversity and stability of herbivorous fishes on coral reefs.
\newblock Ecology 93:891--901.

\bibitem[{Thomas(2010)}]{thomas2010}
Thomas, C.~D.
\newblock 2010.
\newblock Climate, climate change and range boundaries.
\newblock Divers. Distrib. 16:488--495.

\bibitem[{Thomas et~al.(2004)Thomas, Cameron, Green, Bakkenes, Beaumont,
  Collingham, Erasmus, de~Siqueira, Grainger, Hannah, Hughes, Huntley, van
  Jaarsveld, Midgley, Miles, Ortega-Huerta, Townsend~Peterson, Phillips, and
  Williams}]{thomas2004}
Thomas, C.~D., A.~Cameron, R.~E. Green, M.~Bakkenes, L.~J. Beaumont, Y.~C.
  Collingham, B.~F.~N. Erasmus, M.~F. de~Siqueira, A.~Grainger, L.~Hannah,
  L.~Hughes, B.~Huntley, A.~S. van Jaarsveld, G.~F. Midgley, L.~Miles, M.~A.
  Ortega-Huerta, A.~Townsend~Peterson, O.~L. Phillips, and S.~E. Williams.
\newblock 2004.
\newblock Extinction risk from climate change.
\newblock Nature 427:145--148.

\bibitem[{Thorson et~al.(2014{\natexlab{{\em a\/}}})Thorson, Jensen, and
  Zipkin}]{thorson2014a}
Thorson, J.~T., O.~P. Jensen, and E.~F. Zipkin.
\newblock 2014{\natexlab{{\em a\/}}}.
\newblock How variable is recruitment for exploited marine fishes? A
  hierarchical model for testing life history theory.
\newblock Can. J. Fish. Aquat. Sci. 71:973--983.

\bibitem[{Thorson et~al.(2014{\natexlab{{\em b\/}}})Thorson, Scheuerell, Buhle,
  and Copeland}]{thorson2014}
Thorson, J.~T., M.~D. Scheuerell, E.~R. Buhle, and T.~Copeland.
\newblock 2014{\natexlab{{\em b\/}}}.
\newblock Spatial variation buffers temporal fluctuations in early juvenile
  survival for an endangered {Pacific} salmon.
\newblock J. Anim. Ecol. 83:157--167.

\bibitem[{Walther et~al.(2002)Walther, Post, Convey, Menzel, Parmesan, Beebee,
  Fromentin, Hoegh-Guldberg, and Bairlein}]{walther2002}
Walther, G.-R., E.~Post, P.~Convey, A.~Menzel, C.~Parmesan, T.~J.~C. Beebee,
  J.-M. Fromentin, O.~Hoegh-Guldberg, and F.~Bairlein.
\newblock 2002.
\newblock Ecological responses to recent climate change.
\newblock Nature 416:389--395.

\bibitem[{Waples et~al.(2009)Waples, Beechie, and Pess}]{waples2009}
Waples, R., T.~Beechie, and G.~R. Pess.
\newblock 2009.
\newblock Evolutionary history, habitat disturbance regimes, and anthropogenic
  changes: what do these mean for resilience of {Pacific} salmon populations?
\newblock Ecol. Soc. 14:3.

\bibitem[{Winfree and Kremen(2009)}]{winfree2009}
Winfree, R., and C.~Kremen.
\newblock 2009.
\newblock Are ecosystem services stabilized by differences among species? {A}
  test using crop pollination.
\newblock Proc. R. Soc. Lond. B 276:229--237.

\bibitem[{Yeakel et~al.(2014)Yeakel, Moore, Guimar{\~a}es, and
  de~Aguiar}]{yeakel2014}
Yeakel, J.~D., J.~W. Moore, P.~R. Guimar{\~a}es, and M.~A.~M. de~Aguiar.
\newblock 2014.
\newblock Synchronisation and stability in river metapopulation networks.
\newblock Ecol. Lett. 17:272--283.

\end{thebibliography}

\clearpage

\section{Supplemental Material}

\noindent
\textsc{Appendix A.} The \texttt{metafolio} \textsf{R} package.

\noindent
\textsc{Appendix B.} Simulation input parameters and default values.

\noindent
\textsc{Appendix C.} An example straying matrix.

\noindent
\textsc{Appendix D.} Sensitivity illustration with alternative parameter values.

\noindent
\textsc{Appendix E.} An illustration of the correlation between populations.

\noindent
\textsc{Appendix F.} Example simulated time series from alternative conservation scenarios.

\clearpage

\section{Figure legends}

%\begin{center}
%\end{center}
\textsc{Fig. 1}. Flow chart of the salmon-metapopulation simulation.

%\clearpage

%\begin{center}
%\includegraphics[width=2.9in]{../examples/thermal-curve-scenarios}
%\end{center}
\bigskip
\noindent
\textsc{Fig. 2}. Different ways of prioritizing thermal-tolerance conservation. Panel a shows thermal-tolerance curves for ten possible populations and panels b--e show different ways of prioritizing four of those populations. The curves describe how productivity varies with temperature for a given population. Some populations thrive at low temperatures (light greys) and some at warm temperatures (dark greys). Some are tolerant to a wider range of environmental conditions (mid greys) but with a lower maximum productivity. The total possible productivity (the area under the curves) is the same for each population.

%\clearpage

%\begin{center}
%\includegraphics[width=4.0in]{../examples/spatial-arma-sim-full}
%\end{center}
\bigskip
\noindent
\textsc{Fig. \ref{f:ts}}. The components of an example metapopulation simulation. We show, from top to bottom, the temperature signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve. The shaded lines indicate populations that thrive at low (light grey) to high (dark grey) temperatures.

%\clearpage

%\begin{center}
%\includegraphics[width=4.5in]{../examples/spatial-mv}
%\end{center}
\bigskip
\noindent
\textsc{Fig. 4}. The importance of preserving thermal-tolerance diversity through spatial conservation strategies. The conservation strategies correspond to figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink). In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations. The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy. We also show example metapopulation abundance time series for the (c, e) short-term and (d, f) long-term environmental-fluctuation scenarios. The thick grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.

%\clearpage

%\begin{center}
%\includegraphics[width=4.5in]{../examples/cons-plans-n}
%\end{center}
\bigskip
\noindent
\textsc{Fig. 5}. The importance of preserving as many populations as possible when we do not know how thermal-tolerance is distributed. In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations. We show metapopulations in which two through 16 populations are conserved. The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy. We also show example metapopulation (c) rate-of-change and (d) abundance time series for the short-term environmental-fluctuation scenario. The thick grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.

\bigskip
\noindent
\textsc{Fig. 6}. Risk-return trade-off in the case where habitat is lost over time through stream flow reduction. The temperature follows both short-term fluctuations and a long-term increase. Thermal tolerance is randomly conserved. Shading indicates conservation plans where two through 16 populations are conserved. (a) Conserving more populations decreases expected variance but also decreases expected growth rate. Dots show simulated metapopulations and contours show 25\% and 75\% quantiles across 500 simulations per strategy. The thick grey line indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate. Also shown are (b) example metapopulation growth rate and (c) abundance time series from the 2 and 16 population scenarios. Regression lines in (b) illustrate a decreasing growth rate through time.

\clearpage

\section{Figures}

\begin{center}
\includegraphics[height=5.5in]{../examples/simulation-diagram3-sansserif-bw}\\
\textsc{Fig.} 1
\clearpage
\includegraphics[width=2.4in]{../examples/thermal-curve-scenarios}\\
\textsc{Fig.} 2
\clearpage
\includegraphics[width=4.2in]{../examples/spatial-arma-sim-full}\\
\textsc{Fig.} 3
\clearpage
\includegraphics[width=4.5in]{../examples/spatial-mv}\\
\textsc{Fig.} 4
\clearpage
\includegraphics[width=4.5in]{../examples/cons-plans-n}\\
\textsc{Fig.} 5
\clearpage
\includegraphics[width=3.0in]{../examples/cons-plans-squeeze}\\
\textsc{Fig.} 6
\clearpage
\end{center}

\end{spacing}

%\setlength{\parskip}{8pt}
%\setlength{\parindent}{0cm}

 \begin{spacing}{1.1}

 %\setcounter{page}{1}
 %\nolinenumbers
 %\include{som}

 \end{spacing}

\end{document}