### 2014-08-04 inst/ms/appendices/Appendix-C.tex

\documentclass[12pt]{article}
\usepackage{graphicx}

\setlength\parindent{0in}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. \textit{Ecological Applications} VOL:pp--pp.}

\bigskip

\textsc{Appendix C.} Sensitivity illustration with alternative parameter values.

\bigskip

\includegraphics[width=3.2in]{Fig-C1}\\
\textsc{Fig.~C1.} The impact of increasing or decreasing various parameter values on metapopulation return abundance. The different coloured lines represent three example salmon populations. The base case represents the base-case values for the short-term environmental fluctuation scenario.

\end{document}