### 2014-08-04 inst/ms/appendices/Appendix-A.tex
Add back 2 appendices

\documentclass[12pt]{article}
\usepackage{graphicx}

\setlength\parindent{0in}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. \textit{Ecological Applications} VOL:pp--pp.}

\bigskip

\textsc{Appendix A.} An example straying matrix.

\bigskip

\includegraphics[width=3.5in]{Fig-A1}\\
\textsc{Fig.~A1.} An example straying matrix. The rows and columns represent different populations (indicated by population number). Dark blue indicates a high rate of straying and light blue indicates a low rate of straying.

\end{document}