### 2014-08-04 inst/ms/appendices/Appendix-D.tex

\documentclass[12pt]{article}
\usepackage{graphicx}

\setlength\parindent{0in}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. \textit{Ecological Applications} VOL:pp--pp.}

\bigskip

\textsc{Appendix D.} Example simulated time series from alternative conservation scenarios.

\bigskip

\clearpage

\includegraphics[width=4.3in]{Fig-D1}\\
\textsc{Fig.~D1.} Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations. This is the same as Fig.~3 but in colour.
\clearpage

\includegraphics[width=4.3in]{Fig-D2}\\
\textsc{Fig.~D2.} Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations.
\clearpage

\includegraphics[width=4.3in]{Fig-D3}\\
\textsc{Fig.~D3.} Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-D4}\\
\textsc{Fig.~D4.} Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-D5}\\
\textsc{Fig.~D5.} \textbf{Two populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.
\clearpage

\includegraphics[width=4.3in]{Fig-D6}\\
\textsc{Fig.~D6.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.
\clearpage

\includegraphics[width=4.3in]{Fig-D7}\\
\textsc{Fig.~D7.} \textbf{Two populations} conserved with random response diversity and \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-D8}\\
\textsc{Fig.~D8.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-D9}\\
\textsc{Fig.~D9.} \textbf{Two populations} conserved with random response diversity and \textbf{long-term declining stream flow}.
\clearpage

\includegraphics[width=4.3in]{Fig-D10}\\
\textsc{Fig.~D10.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term declining stream flow}.

\end{document}